### 2011-11-30 criteria.tex
Added a list of criteria as an independant tex file

%\documentclass[twoside,a4paper]{report}
\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{pdflscape} %% Used for very big table
\usepackage{moreverb}
\usepackage[table]{xcolor}
\usepackage{listings}
\usepackage{paralist}
\usepackage{fancyhdr}
\usepackage[draft]{fixme}
\fxsetup{footnote}
\fxsetup{nomargin}

\newcommand{\ie}{{\textit{i.e.~}}}
\newcommand{\cf}{{\textit{cf~}}}
\newcommand{\eg}{{\textit{e.g.~}}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compared Features}

\begin{itemize}
    \item{\bf What Can be Represented? The Expressiveness Question}

    \begin{itemize}
        \item{\bf Introduction: main logic formalisms}

    [...]A large number of logic formalism do exist, we shall summarize below the most
    relevant for systems actually deployed in current robotic architectures.[...]

        \item{\bf Open World and Close World Assumptions}

    The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumption names a
    modelling choice on the \emph{completeness} of a knowledge domain. In the close
    world assumption, a proposition that can not be proven true is assumed to be
    false (\emph{negation by failure}), while in the open world assumption, a
    proposition may be considered either true, false or unknown. [...]

        \item{\bf Representation of uncertainty}

	Sources of uncertainty for a robot are two-fold: uncertainty
	\emph{intrinsic} to facts (like \emph{``It may rain tomorrow''}),
	uncertainty caused by imperfect perception of the world (\emph{``Is the
	bottle really on the table?''}). Most logics do not account explicitely for
	uncertainty. It must be either relied on specific logics (like Bayesian
	logics) or on extensions of classical logics.

        \item{\bf Representation of time}

	As an agent acting at human-like time scale and dealing with temporal
	concepts (like actions), a robot may want to represent, and possibly to
	reason, about time. Time representation is split into two distinct
	abilities: representing time points (both in the past -- which is roughly
	equivalent to assignment of timestamps to events the robot perceives -- and
	in the future), and representing \emph{passing time} (durations, timespans)
	like in \emph{``the eggs will be cooked in 10 min''}.

	We call a system that do not account for time (\ie that permanently lives
	in present) \emph{atemporal}.

        \item{\bf Context modeling}

	\emph{Knowledge is contextualized information}\fxfatal{Find someone
	respectable how said that :-)}: it is essential for the robot to associate
	the facts it represents to a \emph{context}. The context carries the keys
	for the interpretation of the information. It covers the \emph{domain of
	validity} of the facts, the \emph{common-sense} knowledge required to
	fill the gaps in the representation\fxfatal{give an example}, \fxfatal{What more?}.

        \item{\bf Possible-Worlds and representing what others know}

	Linked to the context representation, but seen from another angle,
	knowledge representation systems may provide explicit ways to model other
	point of view on the world. This ability is often refered as the
	\emph{perspective taking} ability.

        \cite{Levesque2008}, p. 4

        \item{\bf Introspection: knowledge about the knowledge}

        \item{\bf Modifying its knowledge}

        In DL -> always possible to modify the ABox, not always possible to alter the TBox

    \end{itemize}

    \item{\bf Reasoning Techniques}

    \begin{itemize}
            \item{\bf Prediction and projection tasks}

            Levesque~\cite{Levesque2008} distinguish two main tasks, the \emph{projection task} and the \emph{legality task}.

        \begin{itemize}
            \item{\bf Projection task}: determining whether or not some condition while hold after a sequence of actions.
            \item{\bf Legality task}: determining whether a sequence of action can be performed starting in some initial state.
        \end{itemize}

        \item{\bf Reasoning in uncertainty}

        \item{\bf (Non) Monotonic Reasoning}

        \emph{Monotonic reasoning} means that addition of new assertions to a knowledge base
        can only extend the set of assertions that can be inferred, while a
        \emph{non-monotonic} reasoning scheme may lead to retraction of facts.
        [...]

        \item{\bf Presupposition accomodation}

        \emph{Presupposition accomodation} is the ability for the system to
        automatically create a context allowing to make sense of a proposition.[...]

        \item{\bf Learning}
    \end{itemize}

%%%%%%%%%%%%%%%%%
    \item{\bf Acquiring Knowledge}

    \begin{itemize}
        \item{\bf Knowledge acquisition and modalities merging}
        \begin{itemize}
            \item{\bf Perception}
            \item{\bf Interaction}
            \item{\bf External sources (Web, upper ontologies, ...)}
            \item{\bf Learning}
        \end{itemize}

        \item{\bf Grounding/anchoring strategies}

        \item{\bf Ability to automatically create new object instances}
    \end{itemize}

%%%%%%%%%%%%%%%%%
\item{\bf Practical Integration in Robotic Architectures}

    \begin{itemize}
        \item{\bf Lazy evaluation}

        \item{\bf The planning task}

        \item{\bf Integration with executive layers}

        \begin{itemize}
            \item{\bf Language integration}
            \item{\bf Events}
        \end{itemize}

        \item{\bf Monitoring and debugging}

        \item{\bf Is it fast enough? Scalability and responsiveness}
    \end{itemize}

\begin{itemize}
    \item{\bf Underlying knowledge model}

    \begin{itemize}
        \item  Which underlying knowledge (\emph{common-sense}, \emph{upper knowledge}\ldots{})
        \begin{itemize}
            \item  top-down approach?
        \end{itemize}

    \end{itemize}
\end{itemize}
\end{itemize}

\bibliographystyle{ieeetr}
\bibliography{biblio}

\end{document}