### 2012-10-08 krs-survey.tex
Merged from dissertation
-> many changes in many places

\documentclass[a4paper, twocolumn]{article}

\usepackage[utf8]{inputenc}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subfigure}

% typology diagrams
\usepackage{tikz}
\usetikzlibrary{shapes}

%tables
\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{multirow}
\usepackage{varwidth}

\newcommand{\turn}[3][10em]{% \turn[<width>]{<angle>}{<stuff>}
  \rlap{\rotatebox{#2}{\begin{varwidth}[t]{#1}\bfseries#3\end{varwidth}}}%
  }
\usepackage{pdflscape} %% Used for very big table

\usepackage{paralist}

\usepackage[draft]{fixme}
\fxsetup{footnote}
\fxsetup{nomargin}

\graphicspath{{figs/}}

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf~}}}
\newcommand{\eg}{{\textit{e.g.\ }}}

% Macros related to concepts/statements typesetting
\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}
\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}
\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}

\title{Knowledge Representation for Service Robots: Supply and Demand}
\author{Séverin Lemaignan, Moritz Tenorth, Michael Beetz, Rachid Alami}

\begin{document}

\tikzstyle {taxonomy} = [grow=right, level distance = 3cm, sibling distance=7mm]
\tikzstyle {completetaxonomy} = [grow=right]
\tikzstyle {taxon} = [text width=3cm, text centered,inner sep=0pt, outer sep=2pt] %, draw=blue!50,thick]

\maketitle
\tableofcontents

\begin{abstract}

The aim of this contribution is two-fold: we first propose a typology of
knowledge representation systems for service robotics that covers intrinsic
features like expressiveness or reasoning techniques as well as extrinsic
features like knowledge acquisition and integration into pre-existent robotic
architectures. We base the typology on an initial typical scenario of service
robotics.

We then survey the current knowledge representation system in the community to
identify which knowledge-bound cognitive abilities are today well understood
and implemented, and which would require more research efforts.

We hope this article may bring a clearer picture of the challenges of knowledge
representation and manipulation for the robotic research community.

\end{abstract}

\fxnote{Support material: \emph{What is a knowledge representation} by Davis,
Shrobe and Szolovits,
\url{http://groups.csail.mit.edu/medg/ftp/psz/k-rep.html}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sect|intro}

\subsection{Knowledge as the Root of Cognitive Abilities}
\label{sect|cognitive-abilities}

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Foundations of Artificial
Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
wrote together in 1998:

\begin{quotation}

    Central to this effort is to develop an understanding of the relationship
    between the knowledge, the perception, and the action of [\ldots] a robot. The
    sorts of questions we want to be able to answer are

    \begin{itemize}

        \item to execute a program, what information does a robot need to have
        at the outset versus the information that it can acquire \emph{en route}
        by perceptual means?

        \item what does the robot need to know about its environment versus what
        need only be known by the designer?

        \item when should a robot use perception to find out if something is
        true as opposed to reasoning about what it knows was true in the past?

        \item when should the inner workings of an action be available to the
        robot for reasoning and when should the action be considered primitive
        or atomic?

    \end{itemize}

    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.

\end{quotation}

To be answered, those questions have in common one prerequisite: the robot must
be able to manipulate \emph{explicitly} knowledge, and hence, need a way to
represent the knowledge.

This survey focuses on this knowledge representation issue: we aims at first
establishing a comprehensive typology of representational needs for
robotics in the specific context of service robotics, and then at painting the
current landscape of approaches to the knowledge representation problem in the
research community, in order to summarize the major trends and to identify the
possible shortcomings.

We have classified the features of knowledge representation systems (KRS) into
six broad categories: \emph{what can be represented?} or the
\emph{expressiveness} question ; \emph{how things are represented?} ; \emph{what
reasoning techniques are offered?} ; \emph{how knowledge is acquired and
anchored?} ; \emph{how the knowledge is made available and used by the other
components of the robot?} and \emph{which actual knowledge is stored?} or the
\emph{knowledge instantiation} question.

We have identified over ten projects that are either explicitly advertised as
knowledge representation systems or involve explicit knowledge manipulation
within a robot. By knowledge manipulation, we mean symbolic representation of
assertions, be it static statements on the world or spatio-temporal events, and
reasoning on this representation. Prominent features of each of these systems
illustrate how the various features of a knowledge representation system are
currently implemented.

We have also tried, for each project, to make clear \begin{inparaenum} \item
how informations from perception or interaction are turned into knowledge,
\item how, in return, symbolic concept are \emph{anchored} into the robot
sensori-motor space, \item how the knowledge representation system integrates
with the decisional layers of the robot, and how this can improve or ease
robotic control.\end{inparaenum}
\fxfatal{Do we really do it?}

\paragraph{What do we call ``knowledge''?}
\label{sect|on-knowledge}

Since we will discuss at length the concept of knowledge in the context of
robotics in the coming pages, it is useful to make our terminology explicit.

Be it in philosophy, cognitive sciences or computer sciences, reaching an
agreement on a definition of ``knowledge'' seems difficult.

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language, semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the thesis, give a second look to
this distinction.

In our robotic context, we define knowledge as a narrower concept, while
keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
interrelated logical facts that are meaningful to the robot executive
controller}. By \emph{meaningful} we mean that can possibly be interpreted to
lead to a purposeful action. We will see that our main challenge while
designing a cognitive architecture is furthermore to make this knowledge as
\emph{explicit} as possible.

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question. In this thesis,
we will associate data to low-level material like raw sensor output, and
information to uncontextualised symbolic facts.

To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, book)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

\subsection*{Survey Inclusion Criteria}
\label{sect|inclusion-criteria}

Every robotic system has, implicitly or not, some knowledge representation
systems. It may range from a simple state vector to an explicit symbolic
knowledge base.  This survey focuses on the right end of this spectrum:
symbolic systems, suited for abstract reasoning.

Besides, we have decided to restraint the set of systems to those actually
implemented on robots, and used in semantic-rich environments (\ie dynamic,
partially unknown environments with a large range of different entities which
may have interactions). The typical scenario that would involve such robots is
a service robot in a human-friendly environment like a kitchen (we will
describe such a scenario in section~\ref{sect|scenario}).

We have limited ourselves to systems that
\begin{inparaenum} 
    \item  run on \emph{service robot} (that is, robots that interact with 
    objects in a semantic-rich environment primarily designed for humans),
    \item  ground the knowledge in the physical world (physically embedded
    systems able to assess their environment),
    \item  are able to merge different knowledge modalities,
    \item  are able of on-line, dynamic knowledge acquisition and reasoning 
    (\ie not simple static databases).
\end{inparaenum}

These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.

We have also chosen not to include the GOLOG language and its
derivatives~\cite{Levesque1997, Ferrein2008, Gspandl2011} in this survey. While
several implementations on robots, including service robots, do exist, the
focus of this language is on representation and reasoning about actions and
situations, and the link with symbolic, abstract knowledge is not explicit.

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
GLAIR~\cite{Shapiro2009} or {\sc ACT-R} have declarative knowledge
modules~\cite{Derbinsky2010} and have been recently used on service robots (see
{\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are also absent from this
survey because we did not find much references in the literature on knowledge
manipulation and representation applied to real-world robotic scenarii for
these architectures.

A comprehensive reference on (bio-inspired) cognitive architectures is also
available from the BICA Society~\cite{BicaCogArch2011}.

\subsection{Previous work}
\label{sect|evaluation-literature}

The typology we propose is in part based on a comprehensive synthesis of
classifications and analysis found in the literature. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation in the
context of service robotics.

Levesque and Lakemeyer~\cite{Levesque2008} present in their chapter on
Cognitive Robotics several characteristics of knowledge representation systems
for robots, stressing the need of representing the \emph{dynamics of the
world}.  Sensing is included in the knowledge representation via
\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
distinct parallel mental models; action representation and reasoning about
tasks is discussed in the context of \emph{situation calculus}; \emph{open
world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
robot programming and knowledge representation can be related. We integrate
most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
\emph{knowledge processing middleware} as systems supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context dependent processing at many different levels of abstraction''. They
identify six characteristics: the system must be able to \emph{merge
informations} from different, possibly distributed sources; it should support
quantitative as well as qualitative processing of information, it should offer
\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
(which require what we call here \emph{non-monotonicity}) and finally
\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
specification of the processing functionalities'').

Several surveys compare global cognitive architectures \cite{Langley2006,
Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
distinguish nine capabilities: recognition and categorisation, decision making,
perception and situation assessment, prediction and monitoring, planning,
reasoning, execution control, interaction and
learning/remembering/introspection. They also separately identify four
\emph{properties} of a cognitive architecture, that categorise how knowledge is
handled by the architecture: representation of knowledge, organisation of
knowledge, utilisation of knowledge and acquisition and refinement of
knowledge. This categorisation had a notable influence on our typology, and
many of these categories are also present in our proposal.

Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology}
(how the structure of the world is captured in a representation and shared),
the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.

Chong et al.~\cite{Chong2009} also provide a recent review of the main cognitive
architectures, with a focus on eight functions: perception, memory, goals
management, problem solving capabilities, planning, reasoning, learning and
links to neurobiology.

At an even broader scope, several authors from fields that are connected to
robotics have previously listed desirable features of artificial systems aiming
at rich cognitive abilities.

For instance McCarthy recently listed in~\cite{McCarthy2007} the challenges he
identifies on the road to a \emph{human-level AI}.

\begin{itemize}

	\item the ability to \emph{"operate successfully in the common sense
	informatic situation"},

	\item the necessity of relying on mathematical logic, as the most fruitful
	formalism for machine intelligence,

	\item the ability to deal with \emph{approximate concepts and approximate
	theories} (that would include representing them, and reasoning with them),

	\item non-monotonic reasoning,

	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
	extend \emph{on demand} the closed domain of interpretation for a
	given assertion,

	\item the ability to formalise and reason about contexts,

	\item reasoning about events, and in particular, actions,

	\item the capacity of introspection,

	\item and finally, he points the issue of giving computer the right
	heuristics for decision making.

\end{itemize}

Coming from the perspective of natural language processing in situated context,
Roy and Reiter summarise in~\cite{Roy2005} what they see as the main challenges
to be tackled by knowledge representation systems: \emph{cross-modal
representation} systems, association of words with perceptual and action
categories (\emph{grounding}), modeling of \emph{context}, definition of the
right \emph{granularity} of models, integration of \emph{temporal modeling and
planning}, ability to \emph{match past (learned) experiences} with the current
interaction and ability to take into account the \emph{human perspective}.

Knowledge representation systems in robotics are directly affected by these
points, and we indeed integrate them in our typology, in slightly reformulated
ways.

\subsection{Article overview}
\label{sect|overview}

The paper is organized in four main sections.

The next section proposes a prototypical scenario of a robot acting in a rich
semantic environment. We will not survey knowledge representation systems based
on this scenario: we merely use it to ground the selection and illustrate the
set of features that define the dimensions of a knowledge representation system
for robotics.

Section~\ref{sect|features} presents these features in detail, and organize
them into a typology.

Section~\ref{sect|surveyed-systems} then presents each of the systems we have
examined, and underlines their specificities.

We finally summarize in section~\ref{sect|conclusion} the current approaches in
order to identify several research directions that are not sufficiently
addressed by the current state of the art.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Prototypical Scenario}
\label{sect|scenario}

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
	\caption{The {\em Brownie} scenario}
	\label{fig|scenario}
\end{figure*}

The aim of this imaginary scenario (that has not been implemented, neither in
simulation nor on a robot) is to materialise early in this paper the context
and challenges of knowledge representation and manipulation for service, social
and interactive robotics. It underlines the place, the role and the need of
knowledge in a near-future, everyday situation where several robots and humans
co-exist and cooperate. This scenario will also be a source of support examples
for later sections of the thesis. Also, we have emphasised several keywords in
this description: we will come back to them in the next section to
explain them formally and relate them to each other.

We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
Robi and Roba are two service robots, that can freely move and pick objects
around (with possibly different hardware and software architectures, including
different knowledge representation systems: typically, two robots built by two
different companies). They cooperate with a human in a kitchen environment.

The main task of the scenario is the joint realization of a brownie, initiated
by Tom, the human: ``Let's make a brownie for tonight!''.

The scenario is successful {\it 1)} if the task is achieved (the brownie is baked),
{\it 2)} in a reasonable time (typically shorter than what it would have been
required by the human alone), {\it 3)} and the human is happy by the help it received
(he/she did not feel useless, uncomfortable, unsafe).

We voluntarily do not detail the subtasks of the scenario, neither we define
how they are shared amongst agents: our focus is on knowledge needs and flows.

A ``first-order'' analysis of this task leads to a rough partition of the
required \emph{representation} abilities:

\begin{enumerate}

	\item Representation abilities related to the execution of a complex
	spatio-temporal task,

	\item Representation abilities related to cooperation with other agents.

\end{enumerate}

% Representation of a complex task
We can further refine these categories: to prepare and bake a brownie, the
robot first needs to make sense of the term \emph{brownie} itself: what is it?
what is it used for? what is it made of? etc. We call this knowledge
\emph{common-sense knowledge} and the robot must be able not only to represent
it, but also to have access to an initial source (for instance through a
initial set of facts that are made available at startup, or via access to a
Web-based knowledge base like Wikipedia)

Once bound to the action \emph{make}, this should lead the robot to build and
represent a \emph{context}: we are in a scenario involving cooking. The context
enables the robot to retrieve more common-sense knowledge, like that actions
related to cooking often take place in the kitchen, cooking requires
ingredients, utensils and a procedure that may be provided by a recipe.

These last assertions imply several other capabilities: ``cooking often takes
place in the kitchen'' implies that representation of both uncertainty and
likelihood is desirable. The fact that cooking is associated to a place further
implies that the system models locations and is able to attach \emph{thematic
relations} to concepts (here, the likely location of the cooking action).

``cooking requires a procedure and ingredients'' hints about another important
feature closely tied on knowledge manipulation: \emph{reasoning}. The robot can
\emph{infer} that cooking may require a recipe since a procedure and a list of
ingredients are pre-requisites of the cooking action that may be provided by a
recipe.  If we omit the ``may'', this is a typical example of first-order logic
reasoning.  Many other reasoning techniques exist (including probabilistic ones
-- ones able to deal with the ``may''), we shall illustrate some of them later
in this scenario.

We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
robot should be able to store this plan in a way that allows later execution.
The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
must be as well appropriately handled.

To make decision, a robot may also want to \emph{predict} the state of the
world after some action (``if I leave the cake 2h in the oven, it will burn'').
Such ability to project itself in future or, generally speaking, in other
possible state of the world is related to several cognitive ability and
reasoning techniques: \emph{planning}, \emph{projection}, \emph{representation
of possible worlds} and \emph{non-monotonic reasoning}, in addition to
common-sense knowledge and \emph{physics-based} reasoning (that allows for
instance to predict that an egg is likely to break if dropped).

Procedures are in addition often \emph{underspecified}: we can expect the
recipe to provide a cooking duration, but we usually do not expect the recipe
to tell us to first open the oven door, and then put the cake into it, since it
is self-evident that the door must first be opened to put the cake in the oven.
Our cognitive robot should ideally be able to detect and possibly complete such
underspecification.

% Representation feature that enable cooperation
Then, we want our three agents to cooperate. This, in turn, leads to another
set of cognitive abilities.

Cooperation in our scenario can intervene at many places. For instance, an
agent may want to inform another one about the number of eggs that are
necessary for the brownie. This \emph{helping} behaviour makes sense only if
the first agent knows that the recipient agent both needs the information but
does not know it. This in turn requires the robot to be able to model the
knowledge of the other agents: to think \emph{from the perspective} of another
agent (an idea that is related to the availability of a theory of mind, we will
come back to it later on).

Ability to communicate is one important pre-requisite to collaboration.
Communication in general requires the addresser and the addressee to share a
common interpretative framework (a shared common-sense knowledge -- or cultural
background -- and a shared context). In our scenario, the agents are working in
a kitchen. This element of context does not however suffice if, for example,the 
human asks a robot to ``give {[him]} the bowl''. Behind the symbol
``bowl'', which physical entity are we actually talking about? If we want to
talk and act on the world, this so-called \emph{grounding} operation is
essential. It is a bidirectional process: in covers the {\it top-down}
operation (from the symbol to the percept) and the {\it bottom-up} converse
(retrieval or creation of symbols from perception).

A related ability is called \emph{pre-supposition accommodation}: if one of the
agent moves behind another one, with the brownie dough in its arm, and says
``be careful, I'm behind you!'', we want the first agent to be able to
represent both symbolically and geometrically (because, for instance, if the
agent want to move, it must take into account the new obstacle) something that
is not directly perceived.

Also central to cooperation are the notions of \emph{joint intentions} and
\emph{joint goals}: to help the human during the cooking session, the robots
need to track how far they are into the recipe, what is the next step the human
is likely to go for, how tasks are currently split between agents, what action
is currently blocking the procedure, etc. This knowledge should let the robot
identify the intentions of other agents and create accordingly joint goals.
Hence, a knowledge representation system aiming at dealing with cooperative
behaviours is likely to have goal management structures taking explicitly into
account other agents' actions and goals.

In order to effectively share tasks, the robot must also know what it is
capable of: \emph{capability introspection} (both in term of general capability
and of immediate ability) is thus often desirable. It can be extended to
general introspection (like the ability to tell ``who I am'' or ``what do I
think of'') that may be required for the interaction.

Last but not least, our scenario assumes implicitly \emph{natural interaction}
between humans and robots (as shown by the casual style of the order ``Let's
make a brownie!''), and we want to ensure that the
knowledge available to the robot provides efficient support to the natural
language understanding (for instance by adopting models and vocabulary that are
both well suited for machine processing and remain as close as possible to the
humans own structures and vocabulary), and also to \emph{non-verbal forms of
communication}, like gestures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

We now want to organize requirements from this imaginary cooking scenario where
numerous desirable features for a knowledge representation system for robotics
have been identified without any particular order.

This section proposes a more formal typology and a nomenclature of analysis
axis for a such a system. For each feature, we provide a short definition along
with links to relevant literature.

As previously mentioned, we propose six main categories.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    %[edge from parent fork east]
    %every node/.style={fill=black!30,rounded corners},
    %[parent anchor=east,child anchor=west,grow=east,
    %edge from parent/.style={thick,draw}]
    \node [taxon] {Dimensions}
        child {node [taxon] {{\bf F}. Instantiation}}
        child {node [taxon] {{\bf E}. Integration}}
        child {node [taxon] {{\bf D}. Acquisition}}
        child {node [taxon] {{\bf C}. Reasoning}}
        child {node [taxon] {{\bf B}. Representation}}
        child {node [taxon] {{\bf A}. Expressiveness}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Table~\ref{table|contribution-by-systems}, at the end of the article,
summarizes all these features with the main domain of contribution of each
surveyed KRS.

\subsection{Expressiveness: What Can be Represented?}
\label{sect|expressiveness}

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf A. Expressiveness}
            child {node [taxon] {{\bf A.5}. Meta-cognition}}
            child {node [taxon] {{\bf A.4}. Uncertainty}}
            child {node [taxon] {{\bf A.3}. OWA/CWA}}
            child {node [taxon] {{\bf A.2}. Expressive power}}
            child {node [taxon] {{\bf A.1}. Logic formalism}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

This first axis of analysis is its intrinsic expressive power. It answers the
question: what can be possibly represented. When it explicitly exists, the
\emph{language} of representation plays here an obvious role.

\subsubsection{Main logic formalisms}

The main role of a knowledge representation system is to provide an adequate
representation system to formally store facts and concepts that could be informally
described in natural language.

Formal logic aims at providing such a representation system with the added
value of providing a tractable support for inference and reasoning.

Most (but not all) of the systems we have surveyed rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the fundamental inference problem (called
\emph{satisfiability}: is a given logical sentence true in my model?) in a
tractable manner.

A large number of logic formalisms do exist and we briefly present below the most
relevant ones for systems actually deployed in robotic architectures.

\emph{Predicate logic} is the family of logic formalisms the most commonly
found in knowledge representation. It distinguishes itself from the simpler
\emph{propositional logic} by the use of quantification to increase generality.
\emph{First-order logic} (FOL) is the subpart of \emph{predicate logic} where the
objects of \emph{predicates} (or \emph{formulae}) are simple \emph{terms},
while in \emph{higher-order logics}, predicates can be themselves objects of
other predicates.

\emph{Horn clauses} are an important subset of FOL because the satisfiability
of a set of such clauses is a $P$-complete problem (\ie practically tractable).
A Horn clause is a disjunction of literals (a \emph{clause}) with at most one
positive literal: $\neg p \lor \neg q \lor \cdots \lor \neg t \lor u$, which
can also be represented as $(p \land q \land \cdots \land t) \rightarrow u$.
Important logic programming languages like Prolog are based on Horn clauses.

The family of \emph{Description Logics}~\cite{Baader2008} also plays an
important role. It is also a subset of the first-order logic, with some
extensions in second-order logic. Description logics are notable because most
of them are known to be decidable (but not always in a practically tractable
manner). In description logic, axioms are build from \emph{concepts},
\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
W3C OWL-DL standard is a widely-used language to describe domains with the
description logic.

Because description logics have been originally created from the perspective of
a \emph{knowledge representation language} and not a logic language, their
terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
\emph{individual},\ldots) is well-suited to knowledge description.

\emph{Modal logic}, that allows for statement qualification like
\emph{possibility} or \emph{necessity}, have been shown to be closely related
to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
possible worlds and facts like ``the robot knows \emph{that the human knows}
how to read a recipe''.

\fxwarning{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}

\emph{Temporal logic} are designed to represent and manipulate assertions whose
truth value may vary in time. We introduce one of its key idea (the
\emph{fluents}) later in the typology.

One last class of logics that is of particular relevance for robotic
applications is the \emph{probabilistic logics} or \emph{Bayesian logics}.
These logics provide a formal framework to reason on propositions whose truth
or falsity is uncertain. We elaborate below on the representation of uncertainty.

Note that most of these logic formalisms are still active research field on
their own, and practical considerations (especially the availability of
reasoners efficient enough for on-line use on a robot) often constrain the
choice of a logical formalism and a level of expressive power.

\subsubsection{Expressive power}

Logical formalisms each bring a certain level of expressive power. For
instance, the following classical syllogism can not be represented in
propositional logic because of the use of \emph{universal quantification}:

\begin{quote}
\begin{enumerate}
    \item All men are mortal,
    \item Socrates is a man,
    \item Therefore, Socrates is mortal
\end{enumerate}
\end{quote}

However, the following weak version of the syllogism can be represented in
propositional logic:

\begin{quote}
\begin{enumerate}
    \item If Socrates is a man, then Socrates is mortal,
    \item Socrates is a man,
    \item Therefore, Socrates is mortal
\end{enumerate}
\end{quote}

Generally speaking, expressive power comes at the cost of more complex
\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
at section~\ref{sect|reasoning}.} computations, possibly leading to
untractable, if not undecidable (\ie systems where it is proven that a
proposition can not be decided to be true or false) problems.

\begin{figure}
    \centering
    \includegraphics[width=0.3\columnwidth]{expressive_overlap_dl_horn.pdf}
    \caption{expressiveness overlap of Description Logics and logic programs
    based on Horn clauses, taken from~\cite{Grosof2003}}
    \label{fig|overlap_dl_horn}
\end{figure}

Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
mention extensions to description logics based on rule systems that bring closer
the two approaches.

The relationships between expressive power and reasoning complexity that follow
has been extensively studied for Description Logics.
Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
that allows to conveniently explore these relationships and indexes most of the
literature on that subject.

It can be noted that the relation between expressiveness and reasoning
complexity is fragile: for instance, adding the following axiom \stmt{farFrom
disjointProperty near} (that states that two individuals can not be at the same
time near and far from each other) changes the expressiveness power of the ORO
Common-Sense ontology (presented at chapter~\ref{chapt|oroserver}) from
$\mathcal{SHOIQ(D)}$ to $\mathcal{SROIQ(D)}$\footnote{See
appendix~\ref{chapt|dl} for a brief explanation of this notation}: this
seemingly innocuous assertion change the complexity class of the whole
ontology, and the concept satisfiability reasoning problem switches from a {\it
NExpTime-complete} problem to a {\it NExpTime-hard} problem (\ie, \emph{at
least} as hard as the hardest problem in {\it NExpTime}).

This ``instability'' has practical consequences on run-time performances on
the robot because a light alteration of the knowledge structure can lead to
very noticeable performance drops.

\subsubsection{Open world and close world assumptions}
\label{sect|owa-cwa}

The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumptions name a
modelling choice on the \emph{completeness} of a knowledge domain. In the close
world assumption, a proposition that can not be proven true is assumed to be
false (\emph{negation by failure}), while in the open world assumption, a
proposition may be considered either true, false or unknown.

This distinction is important in robotics where the robot may have to manipulate
concepts with only partial knowledge on them. For instance, let us imagine a robot
that sees a bottle on a table, whose bottom is hidden by another object. The
robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
representation system relying on the closed world assumption would then assume
the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
another example of consequences of the CWA/OWA choice on reasoning.

\begin{table}
    \begin{center}
    \begin{tabular}{ll}
    {\bf Action} & {\bf Part involved} \\
    \hline
    {\tt PickSoftly} & hand \\
    {\tt PickAndPlace} & arm, hand \\
    {\tt MoveArm} & arm \\
    \hline
    \end{tabular}
    \end{center}
    \caption{Assuming the question is: \emph{select actions that do not require
    to move the arm}, a CWA reasoner would return {\tt PickSoftly} whereas an
    OWA reasoner would not return anything if the {\tt PickSoftly} action is
    not explicitly said not to involve the arm.}
    \label{table|cwa-owa-example}
\end{table}

The OWL language is specifically known to assume an open world.  Domains
constrained with the closed world assumption lead to more tractable inference
problems, and allow for instance the use of logic languages like Prolog. Thus,
several approaches exists to \emph{locally close} a domain (\cf
Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those ones).

\subsubsection{Representation of uncertainty and likelihood}

Sources of uncertainty for a robot are two-fold: uncertainty \emph{intrinsic}
to facts (like \emph{``It may rain tomorrow''}), uncertainty caused by
imperfect perception of the world (\emph{``Is the bottle really on the
table?''}). Most logics do not account explicitly for uncertainty. It must be
either relied on specific logics (like Bayesian logics) or on extensions of
classical logics.

\fxfatal{Representation of uncertainty: To be completed!}

\subsubsection{Meta-cognition: knowledge on the knowledge}

As stated by Cox and Raja~\cite{Cox2007}, meta-cognition is composed of both
\emph{``meta-level control of cognitive activities and the introspective
monitoring of such activities to evaluate and to explain them"}.

Sloman proposes in~\cite{Sloman2011} a detailed analysis of meta-cognition and
its different aspects in both natural (human) and artificial systems.

A knowledge representation system endowed with \emph{meta-cognition} is not
only able to manipulate knowledge but also to exhibit and manipulate the
structure of its knowledge and the reasoning process. For instance, the ability
to explain a logical inconsistency in a KRS is a meta-cognitive function, as is
the ability to expose and alter the knowledge structure (these two reasoning
techniques have their own entries in the taxonomy, at
section~\ref{sect|reasoning}.

At section~\ref{sect|introspection} below, we discuss the idea of
introspection.  Meta-cognition can be viewed as the technical facet of the
introspection in general.

%%%%%%%%%%
\subsection{How things are represented?}
\label{sect|higher-level-domain-representation}

We do not discuss in this section the general strategies to construct a
knowledge model (they will be presented in
section~\ref{sect|design-strategies}). We focus here on questions that involve
representational challenges (time, space, context) or require specific
cognitive capabilities (theory of mind, introspection, memory).

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B. Representation}
            child {node [taxon] {{\bf B.5}. Memory}}
            child {node [taxon] {{\bf B.4}. Introspection}}
            child {node [taxon] {{\bf B.3}. Modality}}
            child {node [taxon] {{\bf B.2}. Context}}
            child {node [taxon] {{\bf B.1}. Roles}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\subsubsection{Role representations}

This section discusses strategies and approaches to represent in a knowledge
model three important \emph{roles}: spatial relations, time and representation
of actions.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1 Roles}
            child {node [taxon] {{\bf B.1.3}. Actions}}
            child {node [taxon] {{\bf B.1.2}. Time}}
            child {node [taxon] {{\bf B.1.1}. Space}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\paragraph{Representation of space}

Symbolic representation of space is a widely studied topic. In particular, a
large literature corpus is available on spatial ontologies\fxwarning{Which one?}.

Two main classes of spatial relations are usually represented at the symbolic
level: the topology of environments and the placement of physical entities.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1.1 Space}
            child {node [taxon] {{\bf B.1.1.2}. Placement}}
            child {node [taxon] {{\bf B.1.1.1}. Topology}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Topological maps are abstractions of an environment as graphs where nodes
represent places and edges represent connections between places. Because they
are symbolic representation, topological maps allow higher-level reasoning
(such as containment, connectivity, regions) than metric maps.

One important contribution to the building of a coherent representational stack
for space representation is the Spatial Semantic Hierarchy~\cite{Kuipers2000}
introduced by Kuipers. It consists in multiple interacting representations of
space, both qualitative and quantitative, that span from sensor-level
representation to ontologies of places and regions.

Symbolic representation of entities placement can be absolute or relative. The
relation \concept{isOn}, for example, leads to absolute statements: the
validity of the relation is independent of the nature of its subjects and
objects, and is also independent of the observer.

On the contrary, the relation \concept{nextTo} is relative, and depends on the
relative size of the subject and the object. Two houses distant of 2 meters
from each other can be considered as next to each other. Two ants separated by
2 meters are not next to each other.

The relation \concept{leftOf} is another example of relative spatial relation,
this time because it depends on the observer viewpoint.

\fxwarning{\concept{front}/\concept{back} -> entities have or not a concept of front/back.}

Choices must be made in the knowledge representation system to adequately
represent relative spatial relations. Options include the computation of such
relations only on-demand, when the context is known (which viewpoints, etc.) or
storage of these facts in different models, one per agent.

\fxwarning{Find references on that subject -> classification of spatial
relations. Wikipedia?}

\paragraph{Representation of time}
\label{sect|time-representation}

As an agent acting at human-like time scale and dealing with temporal concepts
(like actions), a robot needs to represent and reason about
time. Time representation is split into two distinct abilities: representing
time points (both in the past -- which is roughly equivalent to assignment of
timestamps to events the robot perceives -- and in the future), and
representing passing time (situations, durations, timespans) like in
\emph{``the eggs will be cooked in 10 min''}.

This is usually formalised as a disjunction in time representation between
\emph{events}, that is, any durationless temporal concept, and
\emph{situations}, that is, any temporal concept with a non-zero
duration\footnote{Other definitions of a situation do exist, notably in the
context of \emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque
consider a situation to be an history of actions.}.

Numerous techniques to represent and reason about time have been devised.
Amongst the most significant ones, we can mention Allen's interval
algebra~\cite{Allen1984} and Ghallab's chronicles~\cite{Ghallab1996}.
\fxwarning{We could detail more chronicles and allen algebra...}
The concept of \emph{fluent} also play an important role for time
representation: fluents are properties (or \emph{conditions}) that change over
time, like $sees(agent1, apple, t)$.\fxwarning{A bit short... citation?}

We call a system that does not account for time (\ie that mentally permanently
lives in present) \emph{atemporal}.

\paragraph{Actions}

Events and actions are two temporal concepts that are of particular importance to
robotic systems, as systems that perceive, react and perform in their environment.

While representation of events boils down to label a timestamp, representation
of actions are more complex since not only they deal with durations, but they
also imply semantic interactions between concepts. From a taxonomy point of
view, actions are a particular type of event that normally leads to a situation
corresponding to the action realisation.

\emph{Thematic roles}~\cite{Gruber1965} (also found as \emph{semantic roles} or
\emph{theta roles} in the literature) allow to semantically qualify the
parameters of an action. The recipient of the action, the performer, the object
acted upon, the destination are some example of common thematic roles.
Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
roles proposed by \cite{Aarts1997} (see~\cite{Gutierrez2001} for a comparison
of other sets of thematic roles present in the literature).

\begin{table}
\begin{center}

\begin{tabular}{lp{5cm}}

\toprule
       Role & Meaning \\
\midrule
      Agent & The \emph{doer} or instigator of the action denoted by the predicate. \\
    Patient & The \emph{undergoer} of the action or event denoted by the predicate. \\
      Theme & The entity that is moved by the action or event denoted by the predicate. \\
Experiencer & The living entity that experiences the action or event denoted by the predicate. \\
       Goal & The location or entity in the direction of which something moves. \\
Benefactive & The entity that benefits from the action or event denoted by the predicate. \\
     Source & The location or entity from which something moves. \\
 Instrument & The medium by which the action or event denoted by the predicate is carried out. \\
   Locative & The specification of the place where the action or event denoted by the predicate in situated. \\
\bottomrule

\end{tabular}
\end{center}

\caption{A list of thematic roles, as proposed by Aarts~\cite{Aarts1997}.
Depending on the action, only certain roles are meaningful.}

\label{table|theta-roles}
\end{table}

In the context of robotic, \emph{task} often represents an aggregate of
(atomic) actions. We use the term here as the representation of the abstract
model of an action. Tasks usually specify at least the conditions required to
performed the action and the consequences of the realization of the action, and
are central to the decisional layers of the robot, in particular for the
planning, monitoring and execution control activities. Knowledge representation
systems may thus have specific mechanisms to represent them (and possible, to
reason about them, as presented in section~\ref{sect|planning}).

Plans representation is closely related to action and task representation. It
is specifically discussed at section~\ref{sect|planning}.

\subsubsection{Context modeling}

Knowledge is contextualized information: it is essential for the robot to
associate the facts it represents to a \emph{context}. The context carries the
keys for the interpretation of the information and set a common ground for
interaction (and, in particular, communication~\cite{Jakobson1960}). It
implicitely defines the domain of validity of the facts and carries the
\emph{common-sense} knowledge required to fill the gaps of the knowledge
explicitely shared between the agents.

Context is generally difficult to recognise, represent and reuse because it is
multiform and largely implicit nature. It is also never unique: at a given
moment, several contexts, at different temporal, spatial, social scales,
overlap.

In the current literature in robotics and cognitive architectures, the term
\emph{context} usually simply refers to a set of beliefs that initiate a
representation (and reasoning) frame: in~\cite{Lemaignan2011a}, the robot
creates a context of interaction with a specific human by storing in a separate
model the beliefs of this human and using this knowledge when dialoging with
the human, the reasoning network is reinitialised in the GLAIR architecture
when the hypotheses that defined the current situation are not believed
anymore~\cite{Shapiro2009}.

This acception of \emph{context} is simplistic, and omit the overlapping and
multi-scale aspects of context modeling.

We see context representation as one of the main challenge of knowledge
representatio in general, and we will further discuss the importance and issues
brought by context modeling in the conclusion of the thesis.

\subsubsection{Modality, contingency and theory of mind}
\label{sect|possible-worlds}

Linked to the context representation, but seen from another angle, knowledge
representation systems may support logical \emph{modality}. A knowledge model
is logically modal if is support the concept of \emph{possible worlds}, \ie,
parallel beliefs models (or \emph{interpretations}) that can be independently
accessed.

A \emph{contingent} proposition is defined as neither always true (a tautology)
nor always false (a contradiction) in every possible world: its truth value
depends on the context. In knowledge representation systems for robotics that
support logical modality, interpretations are often initialised with a common
set of initial beliefs (like common-sense knowledge). This initial common
knowledge is hence true in every possible worlds for the robot, and thus does
not belong to its contingent knowledge.

On the contrary, alternative mental models with contingent knowledge may be
used to represent different (possibly hypothetical or even imaginary) views on
the world, from the robot own perspective or context, or from other
perspectives computed by the robot.

The representation of the mental perspective of other agents has a particular
importance in human-robot interaction. It relies first on the ability to
literally \emph{view} the world from a standpoint which is not
\emph{egocentric}. This cognitive ability is referred as \emph{perspective
taking}. Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} define the
psychological grounds of perspective taking, that are themself originated in
Piaget's work on cognitive development (recent studies on infants include
Moll~\cite{Moll2006} for instance). Perspective taking begins to be studied on
robots as well~\cite{Trafton2005, Breazeal2006, Ros2010}.

The idea of a \emph{theory of mind}~\cite{Premack1978} emerges from the
perspective taking ability. It can be defined as the ability for one to
understand and acknowledge that other intelligent agents can have their own
mental state (that includes beliefs, intents, desires, knowledge) that is
possibly different from one's own. The \emph{attention} plays a central to the
development and recognition of a theory of mind~\cite{Baron-Cohen1985,
Leslie2000}.

A notable consequence of having a theory of mind is the representation of
\emph{false beliefs}, \ie, facts that are believed to be true for an agent, but
false for other ones. The \emph{Sally and Ann} experiment~\cite{Leslie2000}
(figure~\ref{fig|sally-ann}) is a classical example of a false belief
situation. In~\ref{fig|sally-ann-final}, Sally thinks the ball is in the beige
box because she did not see Ann moving it. An external observer asked ``Where
will Sally look for the ball?'' would answer ``in the blue box'' without a
theory of mind (\ie, a model of the knowledge of Sally), whereas it would
correctly answer ``In the beige box'' with a theory of mind.

\begin{figure}
        \centering
        \subfigure[Sally puts the ball in the beige box.]{
            \includegraphics[scale=0.3]{sally_ann1.pdf}
        } \hspace{8pt} %
        \subfigure[She goes away.]{
            \includegraphics[scale=0.3]{sally_ann2.pdf}
        } \\
        \subfigure[Ann moves the ball.]{
            \includegraphics[scale=0.3]{sally_ann3.pdf}
        } \hspace{8pt} %
        \subfigure[Where will Sally look for the ball?]{
            \includegraphics[scale=0.3]{sally_ann4.pdf}
            \label{fig|sally-ann-final}
        }
        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken
        from~\cite{Leslie2000}.}

        \label{fig|sally-ann}
\end{figure}

Scassellati~\cite{Scassellati2002} is one of the first to have implemented a
theory of mind on a humanoid robot.

\subsubsection{Self-knowledge: Who am I? What can I do?}
\label{sect|introspection}

\paragraph{Self-knowledge}

\emph{Self-knowledge} is the term used in philosophy and psychology to describe
the knowledge that an individual has, acquires or infers about itself through
its experiences. It answers the question ``What am I like?''.

Introspection is the technical meaning to access to self-knowledge by the
ability to self describe: what are my capabilities, what is my state
(performing some action, idling, etc.), what are my beliefs, what are my
intentions and my plans?

Introspection must be distinguished from meta-cognition: While introspection
may require meta-cognition (for instance to be able to expose its internal
knowledge), it is not always mandatory. The current state of the robot can be
represented as a simple instantiation of a specific category (for instance, if
the robot gives an object to the human, this state could be represented with the
triples \setstmt{robot performs action1,action1 isA Give}.

\paragraph{Modeling of the robot capabilities}

A particularly important aspect of self-knowledge for robots relates to the
description of its own capabilities: which sensors/actuators/computation
services exist and are currently available ?  While at a first level, these
descriptions can be static (\eg the robot has one laser scanner and two arms),
at more advanced levels, the description is updated and reflect the current
(and possibly past and future) state of the robot. Note that these descriptions
may also involve geometric descriptions (a kinematic chain, the pose of a
device, etc.) that may be deported outside of the main knowledge base. Efforts
trying to formalize, maintain and expose the capabilities and state of a robot
are not new (and ground themselves in work and techniques for self-descriptive
remote procedure calls in computing science), but take a renewed importance
with applications for high-level multi-robot cooperation.

Recent works by Kunze et al.~\cite{Kunze2011} seeks at defining a formal
language to represent the capabilities of a robot.

\subsubsection{Memory}
\label{sect|memory}

Memory has been studied at length in the cognitive psychology and
neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
introduce the idea of \emph{short-term} and \emph{long-term} memory,
Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organises
the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
into a hierarchy. Short-term memory is refined with the concept of
\emph{working memory} by Baddeley~\cite{Baddeley2010}
(Figure~\ref{fig|memory_models}).

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
    \caption{Overview of the main types of memories, based
    on~\cite{Atkinson1968, Anderson1976, Tulving1985, Baddeley2010}}

    \label{fig|memory_models}
\end{figure}

It is worth emphasising that if memory is commonly associated to the process of
forgetting facts after a variable amount of \emph{time}, it actually covers
more mechanisms that are relevant to robotics, like selective remembering
triggered by a specific context or reinforcement learning.

Most knowledge representation systems offers some kind of memory as a pool of
facts that are not forgotten by the robot until it is halted (this memory is
often referred as a \emph{working memory}, but with a meaning unrelated to
Baddeley's definition). Some systems may propose persistent storages that allow
the robot knowledge to grow over time, while others may offer a larger range of
memory categories, like short term memory (that lasts for a couple of seconds)
or episodic memory (that allows the robot to selectively remember facts
associated to specific events).

In the larger field of cognitive architectures, the {\sc Soar}
architecture~\cite{Lehman2006} is one of those that tries to reproduce a
human-like memory organisation. The GLAIR cognitive architecture also have a
concept of long term/short term and episodic/semantic memories.

%%%%%%%%%%
\subsection{Reasoning techniques}
\label{sect|reasoning}

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf C. Reasoning}
            child {node [taxon] {{\bf C.10}. Learning}}
            child {node [taxon] {{\bf C.9}. Naive physics}}
            child {node [taxon] {{\bf C.8}. Planning}}
            child {node [taxon] {{\bf C.7}. Prediction and explanation}}
            child {node [taxon] {{\bf C.6}. Presupposition}}
            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
            child {node [taxon] {{\bf C.4}. Uncertainty}}
            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
            child {node [taxon] {{\bf C.1}. Standard reasoning}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\subsubsection{Standard reasoning techniques}
\label{sect|std-reasoning}

We call \emph{standard reasoning techniques} techniques based on logical
inference, using resolution algorithms like \emph{forward chaining},
\emph{backward chaining} or \emph{semantic tableaux}.

Main reasoning problems include \emph{concept satisfiability},
\emph{consistency checking} and \emph{instance checking}.

Concept satisfiability verifies if it is possible to find a non-empty
\emph{interpretation} of a concept (or an expression defining a concept) in the
knowledge model. For instance, the formula \concept{Plant} $\land$
\concept{isRed}, which defines the concept of red plants, is satisfiable in a
model \concept{KB} iff $\exists a, $ \concept{Plant}$(a) \land$
\concept{isRed}$(a)$, \ie if we can find at least one red plant $a$ in our
model.

Checking the consistency of a model is equivalent to checking the
satisfiability of each of the concept defined in the knowledge model.

Instance checking consists in verifying that an individual $a$ is an
interpretation of a concept (or concept expression) $C$ in the knowledge model.
A typical example would be that we are provided with an instance
\concept{object1} and we want to know if this object is a kind of
\concept{Bottle} or \concept{Glass}.

Inferences can also be drawn from other constructs, whose availability depends
on the representation language. OWL, for instance, has constructs for:

\begin{itemize}
    \item class subsumption (to represent inheritance relations)

    \item reasoning on roles properties, including:
        \begin{itemize}
        \item entailments based on roles domain and range (for instance, if the
        domain of the role \concept{thinksOf} is known to be
        \concept{ThinkingAgent}, then \par \concept{thinksOf}$(a, b) \to
        $\concept{ThinkingAgent}$(a)$),

        \item universal, existential and cardinality constraints,

        \item several second-order predicates (inverse, symmetry, transitivity, etc.)

        \end{itemize}

    \item \emph{class restrictions} like: \par 
	\footnotesize 
	\concept{Bottle} $\equiv$ \concept{Artifact} {\bf that} (\concept{hasShape}
	{\bf value} \concept{cylinderShape})\footnote{This example uses the \emph{Manchester
    syntax}, \url{http://www.w3.org/TR/owl2-manchester-syntax/}}
	\normalsize

    \item set operations like: \par 
    \footnotesize 
    \concept{Color} $\equiv \bigcup$ (\concept{blue}, \concept{green},
    \concept{orange}, \concept{black}, \ldots) 
    \normalsize

\end{itemize}

\paragraph{Rule Languages} As mentioned earlier, knowledge models based on
description logics can be extended through rule languages (typically for OWL,
the SWRL language).

An intersection of properties is an example of expression that can only be
represented with rules. For instance:

\footnotesize
\concept{looksAt(?agt, ?obj)} $\land$ \concept{pointsAt(?agt,?obj)}
$\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
\normalsize

\subsubsection{Dynamic instantiation and alteration of the knowledge structure}

The content of a knowledge base is often conveniently divided into a
\emph{structural part} that defines the conceptualisation of a domain in term
of vocabulary and relations between the concepts, and an \emph{instantiation} of
this structure into concrete entities.

The terms TBox and ABox are commonly found to describe these two different types
of statements in ontologies: TBox statements describe a system structure (made,
for example, of a set of classes and properties) whereas the ABox contains
TBox-compliant statements that are asserted in this structure.

Knowledge representation systems allow to modify the ABox, which can be
considered as the dynamic part of the knowledge base. Alteration of the ABox
include the addition or retraction of relations between existing instances,
and the addition or removal of instances. We call the later \emph{dynamic
instantiation}: the capability for a system to create new instances at
run-time. The term dynamic instantiation applies primarily to concrete entities
(typically, a new object is discovered by the robot, a symbolic instance is
created for it), but may also apply to abstract entities (like an instance of
action, of a feeling, etc.).

The knowledge representation system may also allow to alter the TBox. This
requires the underlying reasoning systems to be able to dynamically take into
account structural changes in the knowledge base.

Example of TBox alteration include modification of the taxonomy (like addition
or retraction of a \concept{subClassOf} relation), changes to the asserted
domain or range of a predicate, addition or retraction of rules.

Supporting TBox alteration has notable consequences on the learning
capabilities of the system: teaching general facts (\ie, facts at the level of
whole categories) like ``cars go on roads'' to a robot requires an alteration
of the TBox.

\subsubsection{Lazy evaluation}
\label{sect|lazy-evaluation}

\emph{Lazy evaluation} describes the ability for a KRS to delay active
knowledge acquisition or reasoning operations until the value is actually
needed.

For instance, a system that computes the symbolic relative placement of two
objects only when this fact is required to answer a query would be said to
adopt a lazy evaluation strategy, whereas a system that computes {\it a priori}
such relations (and by consequence, carries out such computation for possibly
all known objects) would be said to use a \emph{strict evaluation} policy.

Lazy evaluation has an immediate impact on efficiency and scalability of the
system, and some problems may even be only tractable with a lazy evaluation
strategy (the relative placements of object is an example of combinatory
explosion in strict evaluation approaches, although this issue could be
mitigated in real use-cases by heuristics that would select a subset of objects
to evaluate).

One downside is that the knowledge base never contains explicitly the complete
set of beliefs of the robot. This limits for instance the ability for the robot
to react to logical conditions that involve facts that are lazily evaluated
(``trigger a callback when object A is behind object B'' would not be triggered
with a purely lazy evaluation strategy for spatial relations).

\subsubsection{Reasoning with uncertainty}

\subsubsection{(Non) monotonic reasoning}

\emph{Monotonic reasoning} means that addition of new assertions to a knowledge base
can only extend the set of assertions that can be inferred, while a
\emph{non-monotonic} reasoning scheme may lead to retraction of facts.
McCarthy coined a famous example to illustrate the need of non-monotonic reasoning:

\begin{quotation}
Consider putting an axiom in a common sense database asserting that birds can
fly. Clearly the axiom must be qualified in some way since penguins, dead birds
and birds whose feet are encased in concrete can't fly. A careful construction
of the axiom might succeed in including the exceptions of penguins and dead
birds, but clearly we can think up as many additional exceptions like birds
with their feet encased in concrete as we like. Formalised non-monotonic
reasoning provides a way of saying that a bird can fly unless there
is an abnormal circumstance and reasoning that only the abnormal circumstances
whose existence follows from the facts being taken into account will be
considered.
\end{quotation}

An important application of non-monotonic reasoning is the representation of
change: for example, to make a brownie, one needs to crack eggs and mix them to the chocolate.
The eggs disappear and are replaced by a dough:

\concept{Egg}$(a) \land $ \concept{Chocolate} $(b) \land $
\concept{MakeDough}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \land \lnot $
\concept{Chocolate}$(b) \land $ \concept{Dough}$(c)$

The insertion of the proposition \concept{MakeDough}$(a, b, c)$ leads to
retraction of other facts. This rule requires non-monotonic reasoning to be
applied.

\emph{Default logic} is one of the formal logics that account for representing
general truth and exceptions to it (for instance, \emph{tomatoes are red, in
general}). However, due to computational complexity of these models (most of
inferences in default logic are known to be $NP$-complete problem), classical
logics and most of the existing reasoners do not allow non-monotonic reasoning.
For instance, the SWRL rule language, usually associated to the OWL-DL ontology
language, does not allow non-monotonic reasoning (only so-called DL-safe rules
are allowed).

\fxwarning{Make clear 'who does not allow non-monotonic-reasoning': logics? rule
languages? reasoner?}

One important exception is the \emph{negation as failure} inference rule, as
implemented by {\sc Prolog} for instance, that allows for non-monotonicity
within the closed world assumption.

\fxwarning{Give here an example of non-monotonic reasoning with Prolog}
\fxnote{Do we mention here Answer Set Programming?}

\subsubsection{Presupposition accommodation}
\label{sect|presupposition-accommodation}

\emph{Presupposition accommodation}~\cite{VonFintel2008} is the ability for a
system to automatically create a context allowing to make sense of a
proposition.

Applied to robotics, we can imagine a human telling a robot ``Please get me the
bottle that is behind you''. If the robot has not yet see what is behind it, it
needs to assume (and represents in its knowledge model) that a undefined bottle
can be found somewhere in the half of space behind it.

A knowledge representation system able to cope with presupposition
accommodation would be able to take into account this (usually under-defined)
information that is not grounded into perception for later inferences.

This ability to imagine a physically state of the world that is not actually
perceived can be seen as the converse of the grounding ability.

Note also that presupposition accommodation implies a bidirectional link of the
symbolic knowledge model with a geometric (or physical) model of the
environment. This article focuses on symbolic knowledge representation systems,
but we shall mention when a KRS explicitly provides support for presupposition
accommodation.

\subsubsection{Prediction, projection, explanation}
\label{sect|prediction-projection}

Levesque~\cite{Levesque2008} distinguishes two main tasks related to reasoning
on actions and consequences of actions, the \emph{projection task} and the
\emph{legality task}.

We call \emph{diagnosis} the converse of the projection task: the ability to
track back the origin of a decision, and \emph{explanation} the more general
ability to explicit a reasoning or a decision.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf C.6 Prediction and Explaination}
            child {node [taxon] {{\bf D.4} Explaination}}
            child {node [taxon] {{\bf D.3} Diagnosis}}
            child {node [taxon] {{\bf D.2} Legality}}
            child {node [taxon] {{\bf D.1} Projection}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\paragraph{Projection task}: determining whether or not some condition while
hold after a sequence of actions. The projection task is a typical
non-monotonic reasoning task, since at each step, the system must add but also
retract beliefs, as defined in the tasks post-conditions.

\paragraph{Legality task}: determining whether a sequence of action can be
performed starting in some initial state.

\paragraph{Diagnosis}: this corresponds to the ability to rewind on past events
in case of failure to provide possible explanation. This can be seen as the
temporal reverse of the projection task. Because of their modelling of
situations as an history of actions, derivatives of the GOLOG logic programming
languages are a good example of the diagnosis task integrated to the knowledge
representation system~\cite{Gspandl2011}.

\paragraph{Explanation} Diagnosis is also linked to the \emph{explanation} or
\emph{justification} capabilities that may be offered by the knowledge
representation system. The explanation of an entailment is the sequence (or set
of sequences if several are possible) of reasoning steps that allow to reach a
conclusion. An explanation can also conversely explain why a statement leads to
a contradiction.

The following example shows an explanation for an inconsistency in a particular
knowledge base: by adding the statement \stmt{robot1 belongsTo human1}, we
observe that an inconsistency is triggered. The reasoner provides the four
following observations to explain the inconsistency:

\begin{quote}
\scriptsize
\begin{enumerate}
    \item \concept{robot1} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
    \item \concept{belongsTo} {\bf Domain} \concept{Artifact}
    \item \concept{robot1} {\bf Type} \concept{Agent}
    \item {\bf DisjointClasses:} \concept{Agent}, \concept{Artifact}
\end{enumerate}
\normalsize
\end{quote}

This explanation directly portrays the underlying structure of the knowledge
model: we can understand that a robot is modelled here as an agent, that agents
and artifacts are disjoint classes and that only artifacts can belong to
someone, hence the inconsistency. A KRS may provide mechanisms to expose this
kind of analysis, either automatically when inconsistencies occur, or on
demand.

Explanation of contradictions plays a particular role for robots: from a
cognitive point of view, a logical inconsistency (\ie, a contradiction) can be
viewed as a \emph{cognitive conflict} or \emph{cognitive dissonance} (\ie, two
incompatible models of the world that must be dealt with). Being able to expose
and explain such cognitive conflicts eases the control of the behaviour of the
robot in unexpected semantic situations, and forms a first step towards an
adequate reaction.

Cognitive dissonance is also identified by developmental psychologists like
Piaget as a motivation factor for a child to progress through the various
stages of its cognitive development. This has been also studied in
robotics~\cite{Oudeyer2007}.

The cognitive capability of \emph{justification} is also intimately linked to
the meta-cognition capability and participates to the overall \emph{cognitive
observability} of the system.

\subsubsection{Task planning}
\label{sect|planning}

Symbolic task planning is the ability for a robot to select a sequence of
actions in order to reach a given final state. This capability is closely
related to the previously mentioned projection and legality tasks: for a robot
to plan, it must be able to build hypothetical states of the world that would
follow from the successive application of actions (prediction task), and at
each step, select possible, legal actions based on their pre-conditions
(legality task). As already mentioned, these tasks are highly non-monotonic.

Symbolic task planning in general is a large research
field~\cite{Russell2009planning}. The so-called Classical Planning Problem,
first, is characterised by a unique known initial state, durationless
deterministic actions which can be taken only one at a time, and a single
agent. STRIPS and PDDL are amongst the commonly used languages for representing
such planning problems.

Planning with nondeterministic durationless actions with probabilities can be
represented as discrete-time \emph{Markov decision processes} (MDP), and when
full observability is replaced by partial observability, we deal with
\emph{partially observable Markov decision process} (POMDP).

\emph{Hierarchical Task Networks} (HTN) are another common formalism for
planning problems, where an initial set of tasks (\emph{High Level Tasks}, HLA)
is decomposed into either primitive actions or a new set of subtasks.

From the observation that the core reasoning techniques (back and forward
chaining) are shared between planners and reasoners used in knowledge
representation systems, task planning can be considered within the KRS.

Since McCarthy's \emph{Situation Calculus} in 1963, numerous knowledge
representation formalisms dedicated to representation and reasoning about
actions and situation have emerged: besides situation calculus, \emph{fluent
calculus} and \emph{event calculus} are the main ones.
Thielscher~\cite{Thielscher2011} recently proposed a unification of these
approach in a new \emph{action calculus}.

The GOLOG~\cite{Levesque1997} language and its derivatives (like {\sc
ReadyLOG}~\cite{Ferrein2008} and {\sc IndiLOG}~\cite{Gspandl2011}) propose
implementations of the situation calculus focused on robotic applications.

It is however also common to rely on external components dedicated to symbolic task
planning (in particular because of the non-monotonicity requirements) with a
tight link to the KRS for domain retrieval and/or resulting plan storage.

Finally, in the context of interaction between several agents, the management
of \emph{joint intentions} and \emph{joint goals}~\cite{Tomasello2005,
Dominey2011} are additional aspects that have to be represented and
appropriately handled by the planning subsystem.

%An extensive report on task modeling in OWL ontologies in appendix

\subsubsection{Physics-based reasoning}
\label{sect|physics}

As embodied entities, robots have to interact with physical entities.
\emph{Naive physics reasoning} covers all the everyday reasoning the humans
unconsciously perform, like taking into account gravity (``if I drop a ball, it
falls down'') or common physical properties of objects (``a glass may break if
dropped'', etc.). Many of the interactions with our everyday environments are
ruled by such laws that are difficult to exhaustively encode.

Some systems \cite{Kunze2011a} rely on external dedicated physics engine to
compute symbolic facts from on-demand physics simulation. This requires a tight
integration between the symbolic model and a geometric model that carries the
geometries and physical properties of objects.

\subsubsection{Learning}
\label{sect|learning}

\fxfatal{Learning section to write}

%%%%%%%%%%%%%%%%%
\subsection{Acquiring knowledge}

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf D. Acquisition}
            child {node [taxon] {{\bf D.3} Motivation}}
            child {node [taxon] {{\bf D.2} Grounding}}
            child {node [taxon] {{\bf D.1} Acquisition and fusion}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\subsubsection{Knowledge acquisition and modalities fusion}
\label{sect|knowledge-acquisition}

In the survey inclusion criteria, we have chosen to only consider robotic
systems that acquire knowledge by themselves, at runtime.

In our context, \emph{acquiring knowledge} means to build new logical
statements from data sources and to anchor them into the existing knowledge. We
consider three possible sources of data: proprioceptive/exteroceptive sensing,
interaction with other agents, humans or robots, and remote knowledge bases.
The acquisition process has generally two steps: the information acquisition by
itself, and the \emph{transformation} of the information into knowledge,
\emph{aligned} with the robot existing model (following our terminology for
information and knowledge, as discussed in the introduction).

It must be observed that knowledge acquisition is generally not done directly
in the knowledge representation system. External components (often aggregated
into knowledge acquisition \emph{pipelines}) are usually required to convert
percepts into symbolic facts and to ground them.

While we do not review in this article all these systems, the whole process of
knowledge acquisition is central in cognitive robotic architecture and the
design of knowledge representation systems can influence or be influenced by
the approach to knowledge acquisition.

In particular, complex robotic systems often require multi-modal perception
capabilities (for instance, a robot can only interpret an utterance like ``this
is a plate'' if it is able to understand gestures, understand natural language
and merge them in a timely manner). Multi-modal interpretation can take place
at various levels, but in many cases (especially if the modalities are of very
different natures, like in the example above) merging will require
symbolic-level reasoning. The KRS has a direct impact on the feasibility and
ease of such operations.

Let review shortly the three sub-categories of knowledge acquisition.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf D.1 Acquisition and fusion}
            child {node [taxon] {{\bf D.1.3} Linked Resources}}
            child {node [taxon] {{\bf D.1.2} Interaction}}
            child {node [taxon] {{\bf D.1.1} Sensing}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\paragraph{Sensing}

From the point of view of knowledge representation, the sensing capability can
be split into proprioceptive sensing (\ie, sensing of the robot own internal
state) and exteroceptive sensing (sensing of the robot environment). The
(physical) introspection capabilities of the robot relies on the former.

Exteroceptive sensing is the most obvious and largely studied mean of knowledge
acquisition. Traditional sensing devices (IR, cameras, laser-scanners), while
still present and widely used (navigation based on 2D localisation and obstacle
avoidance is today the standard), are being step-by-step replaced by
\emph{synthetic sensors}.

These synthetic sensors include post-processing to provide higher-level
percepts that ease the grounding. The prototypical example of such a device is
the Kinect sensor. At a first level, it replaces traditional stereo vision
algorithms by providing a fast, robust depth map. At a second level, it
provides accurate, real-time tracking recognition and tracking of whole body
poses of human.

Other examples of such synthetic sensors exist: face recognition, off-the-shelf
performant SLAM solutions, automatic cluster segmentation in point clouds (with
the PCL library), etc.

While the progresses of these sensing technologies are remarkable, one field of
perception remain a very difficult challenge: accurate and generic object
recognition. Most of the current approaches to object recognition rely on a mix
of point cloud segmentation and fitting with visual feature recognition
(SIFT-like algorithms), but it remains a slow and fragile task.

\paragraph{Interaction}

Interaction with other intelligent agents (humans or robots) is another
important source of knowledge acquisition. It relies obviously on some form of
sensing (from speech recognition to gesture recognition) but we distinguish it
from the previous section because \emph{interaction} implies a form of
communication. Communication is associated to specific functions
(as shown on Jakobson's diagram, figure~\ref{fig|communication-model}), and in particular, it implies a shared
\emph{context} (usually implicit) between interactors.

\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{jakobson_communication_model.pdf}

    \caption{The Jakobson~\ref{Jakobson1960} communication model. In italics,
    the associated \emph{communication functions}.}

    \label{fig|communication-model}
\end{figure}

We distinguish between two main interaction channels: verbal and deictic.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf D.1.2 Interaction}
            child {node [taxon] {{\bf D.1.2.2} Deictic Interaction}}
            child {node [taxon] {{\bf D.1.2.1} Verbal Interaction}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\label{sect|nlp}

The field of verbal interaction processing for robots spans from pattern-based,
constrained sentences recognition to natural, bidirectional, unconstrained
verbal communication. A large literature corpus exists on Natural Language
Processing (NLP) which is presented at section~\ref{sect|dialogs-related-work},
page~\pageref{sect|dialogs-related-work}.

NLP is an established research field by itself, and while the robotic community
is still lagging behind on many theoretical aspects, it brings one important
aspect: the embodiment. Because the interactors, both the robot and the human,
are establishing a communication within a shared physical context, the verbal
communication channel is complemented by deictic channels, back channels and
possibly shared physical experiences: a human can show something to a robot,
saying ``Give me this''. This is not possible for a virtual agent.

Several of the knowledge representation systems we present have developed
specific built-in mechanisms or extensions to parse, ground and possibly
rebuild natural language.

Deictic (used in the literal meaning of ``display, demonstration, reference'')
interaction is also an established field of research in human-robot
interaction. Common deictic forms of communication~\cite{Li2012} include
attentional focus (via face and gaze tracking) and joint attention, pointing,
emotional expressions (based on face expressions, postures, emotional
gestures).

Like other knowledge acquisition modalities, the recognition and interpretation
of deictic communication is rarely directly included in a KRS, but, as
previously mentioned, the symbolic representation of such communication acts is
relevant and important to achieve successful multi-modal interactions.

\paragraph{Linked Knowledge Resources}
\label{sect|lod}

Robots, and in particular service robots, have usually an access (with possibly
security-related constraints) to the World Wide Web and remote knowledge stores.

The current shift towards the Semantic Web (\ie structured, annotated data that
are easily machine-processable) makes increasingly easy to have robots to reuse
autonomously this knowledge. The DBPedia project, for instance, illustrates
well the tendency: it provides an automatically generated RDF version of the
Wikipedia encyclopedia.

In its current state, the relevance of the DBPedia project in our context is
however limited: the triples that are extracted are mostly ``mechanical'' (like
the categories of a term, or factual informations extracted from Wikipedia's
InfoBox), and the vast majority of the knowledge actually contained in the
encyclopedia pages remains out of reach of automated parsers. Efforts in that
direction however exist~\cite{Nyga2009, Fader2011}.

Another notable project that seeks at providing large amount of
machine-friendly common-sense knowledge is the MIT's OpenMind
project~\cite{Singh2002}. The project is designed to let the general public
easily add common-sense statements in semi-controlled natural language, which is
then processed to a publicly available ontology.

Another approach that is easier at short-term to let robots remotely access
knowledge repositories consists is building robot-specific shared repositories,
with declarative and/or procedural (\ie, plan library) knowledge. The
RoboEarth~\cite{Waibel2011} project is an example of such an effort.

\subsubsection{Grounding/anchoring strategies}
\label{sect|grounding}

\emph{Grounding} (also called \emph{anchoring} when specifically referring to the
building of links between \emph{percepts} and \emph{physical
objects}~\cite{Coradeschi2003}) is the task consisting in building and
maintaining a bi-directional link between sub-symbolic representations (sensors
data, low-level actuation) and symbolic representations that can be
manipulated and reasoned about~\cite{Harnad1990}.

Being embodied entities with interaction with other embodied entities as a
fundamental requirement, robots and robotic is deeply concerned by the
grounding issue.

Being actually implemented on real service robots, all the symbolic knowledge
representation systems that we review in this study have some kind of grounding
process. Numerous approaches exist, like amodal
\emph{proxies}~\cite{Jacobsson2008}, grounded amodal
representations~\cite{Alami2011, Mavridis2006}, semantic maps
(Figure~\ref{fig|semanticmap}, \cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planing and object classification~\cite{Lorken2008,
Varadarajan2011}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{semanticmaps_hertzberg.png}
    \caption{Example of a semantic map, taken from~\cite{Nuechter2008}.}
    \label{fig|semanticmap}
\end{figure}

\subsubsection{Intrinsic motivation and curiosity}

The reasons for a robot to acquire knowledge are diverse, and usually external
to the robot itself. It is often driven by the requirements of tasks that the
robot has to execute. In this case, motivations are managed by the execution
controller and are mostly invisible to the knowledge representation system.

However, motivation can also be intrinsic, driven by the internal state of the
robot's beliefs, without external pressure or reward. In this case, the \fxwarning{to be completed}

In the context of robotics, Oudeyer notes however that \emph{the information that is
compared \emph{[to compute a level of motivation]} has to be understood in an
information theoretic perspective, in which what is considered is the intrinsic
mathematical structure of the values of stimuli, independently of their
meaning.}~\cite{Oudeyer2007}.

Psychological grounds of motivation are summarized in~\cite{Oudeyer2007} while
the main approaches to computational motivation, divided into knowledge based
models, competence based models and morphological models, are surveyed
in~\cite{Oudeyer2008}.

%%%%%%%%%%%%%%%%%
\subsection{Practical integration in robotic architectures}
\label{sect|integration-robot}

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf E. Integration}
            child {node [taxon] {{\bf E.4} Performances}}
            child {node [taxon] {{\bf E.3} Monitoring}}
            child {node [taxon] {{\bf E.2} Executive layers}}
            child {node [taxon] {{\bf E.1} Sensori-motor}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Knowledge representation systems do not mean anything to robots if they are
considered in isolation. This section proposes categories of features related
to the integration of the KRS into a larger software architecture that includes
perception routines, decision-making processes and actuation control.

We also mention some practical aspects of a real-world system, like
performances and monitoring tools that come along with the KRS.

\subsubsection{Integration with sensori-motor layers}
\label{sect|integration-sensorimotor}

We have previously discussed (section~\ref{sect|grounding}) the principles of
the grounding process that aims at establishing and maintaining a connection
between percepts (and to a lesser extend, low-level actions) and symbols.

While every real-world cognitive robot need some kind of grounding, the actual
implementations lead to very different information flows.

The systems can be roughly split into two classes: \emph{passive} knowledge
repositories that process symbolic facts produced by lower-level sensori-motor
layers (\emph{push} flow); \emph{active} knowledge managers that directly query
(possibly by polling or on-demand) low-level layers.

This macroscopic distinction is however mostly a matter of defining the
frontiers of the KRS: some systems like KnowRob~\cite{Tenorth2009a} encompass
geometric reasoning layers that would be considered as external by other
systems like ORO~\cite{Lemaignan2010} that focus on the symbolic fact storage
and rely on a ecosystem of independent modules to provide and consume symbolic
knowledge.

\fxwarning{Some archi with the ability to ``listen'' to the robot internal
structures -> which ones?}

Other systems do not fit either in such a partition between active and passive
systems because they do not stand as independent modules but exist as diffuse,
\emph{ubiquitous} knowledge manipulation system (case of the CAST knowledge
model~\cite{Jacobsson2008}), for instance because they are primarily
language~\cite{Ferrein2008, Sabri2011}.

\subsubsection{Integration with executive layers}
\label{sect|integration-executive-layers}

Conversely, the knowledge management module need a tight integration with the
decision-making processes. As for the integration with sensori-motor
layers, the borders of the KRS can be fuzzy and vary from one architecture to
another: many consider symbolic task planning as an integral role of the KRS,
while other have dedicated extensions for planning, some integrate learning as
an on-the-flight process that is part of the KRS, others as an independent
deliberative process, etc.

The actual integration techniques vary also widely, from language extensions
(like the integration of CRAM~\cite{Beetz2010} with KnowRob) and client-server
architectures, to event-driven models (SHARY and ORO~\cite{Alami2011}). Choices
at this level have notable consequences on the whole design of the upper
control architecture of the robot, in particular regarding its modularity and
the ease of addition of new components.

\subsubsection{Monitoring and debugging}
\label{sect|debugging}

It is common to have knowledge representation systems at the heart of a
cognitive robotic architecture, and therefore KRS are easily ``buried'' in the
system.

At the same time, the symbolic model often provides a valuable synthetic view
on the whole state of the robot, furthermore easily understandable by the human
developer (the fact \stmt{human1 isSitting true} is easier to interpret than
the suite of relative coordinates of each joints of the human skeleton, as
provided by the human tracker, for instance).

Tools to trace and visualise at run-time the evolution of the knowledge
structure and contents may be available with the KRS, as well as
post-processing tools that run on the trace to analyze {\it a posteriori} the
cognitive behaviour of the robot.

\subsubsection{Evaluation of performances}
\label{sect|performances}

Benchmarks of symbolic systems for robots are hard to conduct for several
reasons: identifying good metrics for robotic experiments in general is
difficult because of the complex interactions between tenth of modules running
in parallel, and isolating one specific component is difficult.  Also,
knowledge representation systems are often tightly coupled to the other
modules. To quote Langley~\cite{Langley2006}:

\begin{quote}
The conventional wisdom of software engineering is that one should
develop independent modules that have minimal interaction. In contrast, a
cognitive architecture offers a \emph{unified} theory of cognition with tightly
interleaved modules that support synergistic effects.
\end{quote}

The lack of standard API for knowledge services makes it also hard to switch
between KRS to compare them.

Finally, because service robots are designed to act in rich, dynamic
environments, possibly with humans, building repeatable experiments is
challenging, and quantitative measurements are often not the right
metric~\cite{Langley2006}.

We will however present here some quantitative metrics (related to scalability,
for instance), followed by qualitative evaluation approaches, grouped under the
term \emph{Cognitive Performances}.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf E.4 Performance \\Evaluation}
            child {node [taxon] {{\bf E.4.2} Cognitive Performances}}
            child {node [taxon] {{\bf E.4.1} Raw Performances}};
 \end{tikzpicture}
\end{center}
\end{scriptsize}

\paragraph{Raw Performances} The \emph{raw performance} is evaluated on
quantitative benchmarks. The main metric is the \emph{scalability}
$\mathcal{S}^{KRS}$ of the system with the size of the knowledge base
$\sigma=|\Delta^{\mathcal{I}}|$  (in term of atoms or statements).

We call \emph{relaxation time} $\mathcal{R}^{\mathcal{M}}$ the (averaged) time
required by the system after a model modification of type $\mathcal{M}$ before
being available for further interaction, and \emph{query time}
$\mathcal{Q}^{\mathcal{M}'}$ the (averaged) time to execute a query of
complexity $\mathcal{C}$ on the KRS. The type $\mathcal{M}$ of model
alteration is either an ABox modification (addition/removal of an instance) or
a TBox alteration (addition/removal of a class, a class restriction or a rule).
%Query complexity $\mathcal{C}$ is...\fxwarning{find references}.

\emph{Temporal scalability} is defined in term of the nature of the function
$f^{\mathcal{M}, \mathcal{C}}(\sigma) = \mathcal{R}^{\mathcal{M}}(\sigma) +
\mathcal{Q}^{\mathcal{C}}(\sigma)$ (\ie the relation of the relaxation time
and query time to the knowledge base size). \emph{Space scalability} is the
relation of memory consumption to the size $\sigma$. The scalability is tightly
coupled to the expressiveness of the underlying knowledge model, which need to
be known for the scalability measurement to be meaningful.

Because of coupling and repeatability issues we have mentioned, raw
performances of KRS are often benchmarked with synthetic datasets (which leads
to another issues: how to assess the meaningfulness of the performance of a
reasoner on an artificial ontology?~\cite{Bail2010}) or ``toy'' experiments
that do not always model the whole complexity of real-world
application~\cite{Chong2009}.

\paragraph{Cognitive Performances} While evaluating the raw performances of
knowledge representation systems in a relevant manner may be difficult, the cognitive
performances of the robot as a whole can be also evaluated.

Langley et al.~\cite{Langley2006} propose five such dimensions of evaluation:
the \emph{generality} of the system (can it adapt easily to new tasks?), the
\emph{rationality} or relevant of the inference/reasoning/decisions the system
take, the \emph{reactivity} and \emph{persistence} that evaluates if the
behaviour of a cognitive system is appropriate under unpredicted changes, the
\emph{improvability} of the system as a function of the knowledge added to it,
and finally, the resulting \emph{autonomy} of the system.

Cognitive performance can also be evaluated with the support of tools developed
in cognitive psychology. Several standard tests (like False-Belief
experiments~\cite{Leslie2000} or the Token test~\cite{DiSimoni1978}) have been
used to judge the cognitive abilities of robots~\cite{Mavridis2006,
Breazeal2006}.
%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Knowledge instantiation}

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf F. Knowledge instantiation}
            child {node [taxon] {{\bf F.4} Granularity}}
            child {node [taxon] {{\bf F.3} Metrics}}
            child {node [taxon] {{\bf F.2} Common-sense and Alignement}}
            child {node [taxon] {{\bf F.1} Design Strategy}};
 \end{tikzpicture}
\end{center}
\end{scriptsize}

This last branch of our taxonomy looks at the actual \emph{content} of the
knowledge base: the knowledge instantiation. Here, \emph{instantiation} does
not only refer to the instantiation of the knowledge structure (what we have
called the ABox), but also includes the knowledge structure itself (the TBox).

While we have previously mentioned features of knowledge representation systems
that enable the robot to fill its knowledge base with content and alter the
knowledge structure, most of the systems also come with a certain amount of
initial knowledge that often includes \emph{common-sense} knowledge (\ie facts
that widely known to humans, and hence often implicit: ``to put a cake in the
oven, one must first open the oven's door'').

The design strategy, the choice to rely on common-sense knowledge or not, the
reuse of standard ontologies, the quantity of {\it a priori} knowledge are many
parameters that lead to different knowledge models.

\subsubsection{Design strategy}
\label{sect|design-strategies}

The main challenge of knowledge representation can be summarised as \emph{How
to model the real-world state and interactions in a symbolic way, processable
by the robot to make decisions}. We have already introduced the term
\emph{grounding} to describe the (bi-directional) process of binding percepts
to symbols.

We have also seen that the instantiation of the knowledge (\ie, the actual,
practical knowledge available to the robot) comes either from some variant of
perception plus grounding, or from knowledge that the developer considers as
already meaningful for the robot: remote semantic databases or initial
(common-sense or situation specific) knowledge.

This translates into two main strategies to drive knowledge instantiation: a
\emph{top-down} design or a \emph{bottom-up} design.

A bottom-up approach to knowledge instantiation takes the output of sensors as
the primary source of knowledge: instances of objects, agents, and their
relations directly result from what is perceived. No abstract, ``from the more
generic to the more specific'' process takes place.

Steels~\cite{Steels2007} considers this approach to be a solved issue, and
according to Sloman~\cite{Sloman2007} (and his stance against the
\emph{``Symbol Grounding meme''}), since bottom-up grounding boils down to grounding
of \emph{somatic} concepts (\ie roughly, the sensori-motor relationships that
the robot learns from its interaction with the world), it constrains in an
unacceptable way the range of concepts accessible to the robot.

Knowledge instantiation can also be approached as a \emph{top-down} activity:
in natural language grounding, for instance, the robot needs to automatically
bind an abstract representation (a group of words uttered by a human) to an
unambiguous, context-dependent, internal concept. This concept may (or may not)
be \textit{a priori} available to the robot as a pre-loaded ontology (what we
previously called the cultural background of the robot). In any case, the robot
must conduct a cognitive process that leads from an abstract concept to a
concrete entity.

The bottom-up and top-down strategies also reflect how the knowledge structure
itself is constructed: either by successive classification and refinement of
percepts, or from generic categories, typically extracted from standard
upper-ontologies.

\subsubsection{Common-sense and alignment with standard upper-ontologies}

At section~\ref{sect|lod} we have presented how remote knowledge bases are a
valuable source of knowledge, including common-sense knowledge, that robot can
extract.

Building {\it a priori} knowledge with a top-down strategy leads to populate
the upper part of the taxonomy with abstract concepts like \emph{Thing},
\emph{Time}, \emph{Person}, etc. The organisation of the whole knowledge
depends on the design of this abstract part of the common-sense knowledge.

Different knowledge structures at this level can lead to serious
misunderstanding: for instance, if one robot considers the concept of a
\emph{person} as representing some intelligent, possibly disembodied, entity,
and another robot represents a \emph{person} as a subclass of mammals, their
models of the world are likely to be conflicting.

To be able to successfully exchange knowledge with other systems (robots,
databases, natural language parsers, etc.), the common-sense knowledge of
robots is thus often aligned with a standard upper-ontology.

Many such upper-ontologies exist (table~\ref{table|upper_onto_metrics} lists
some of them). While {\sc Cyc} and SUMO are current the main two, many efforts
take place to make these ontologies compatible with each other (in particular
by using the {\sc WordNet} thesaurus as intermediate unambiguous source of
semantics).

\subsubsection{Metrics and quality criteria}

Qualitative and quantitative metrics give an insight on the size, complexity
and effectiveness of ontologies used with the robots.

Table~\ref{table|upper_onto_metrics} gives such metrics for five major
upper-ontologies commonly used in the semantic Web community (taken
from~\cite{Mascardi2007} and the projects' respective websites). It is
interesting to note the large variations between projects like {\sc DBpedia}
whose content is automatically generated from Wikipedia, {\sc Cyc} or {\sc
SUMO}, which are both hand written, but do not adopt the same strategy to
select the knowledge to represent.

These metrics do not reflect adequately the \emph{expressive complexity},
though: in most of these ontologies with a large amount of terms and
assertions, taxonomic relations (\concept{isA}) or technical predicates (URI,
translations, etc.) account for a large part of the assertions, at the expense
of real semantic relations.

\begin{table}
\begin{center}

\begin{tabular}{llllll}
\toprule
{\bf Project} & {\bf Terms} & {\bf Assertions} (triples) \\
\midrule
{\sc Cyc} & > 300 000 & > 3 000 000 & \\
YAGO & > 10 000 000 & > 120 000 000 \\
SUMO & 20 000 & 60 000 \\
{\sc DBpedia} (for English) & 1 840 000 & 385 000 000\\
{\sc OpenMind} Common Sense (for English) & & 1 000 000\\
\bottomrule

\end{tabular}
\end{center}
\caption{Raw size of major upper-ontologies.}
\label{table|upper_onto_metrics}
\end{table}

Other metrics that include the type of predicates that are used, or the
computed DL expressiveness, can be more significant (some are presented in
table~\ref{table|onto-stats}, page~\pageref{table|onto-stats}).

Qualitative evaluation of ontologies is also a well studied field. In Staab's
{\it Handbook on Ontologies}, Vrande{\v{c}}i{\'c}~\cite{Vrandevcic2009}
provides a synthesis of qualitative criteria for ontology assessment.  He lists
eight of them: \emph{accuracy, adaptability, clarity, completeness,
computational efficiency, conciseness, consistency,} and
\emph{organizational fitness}. Details and relevant literature can be found
in Vrande{\v{c}}i{\'c}'s chapter.

\subsubsection{Granularity}

Amongst the characteristics of ontologies, knowledge \emph{granularity}
qualifies the level of details or refinement of the knowledge stored. The level
of granularity of a robot's ontology hints on the place of the symbolic layer
in the whole robotic architecture: some systems (like OMRKF~\cite{Suh2007}) go
as down as storing SIFT features (\ie a large volume of numerical values) in
the ontology. The storage of literal values is indeed a relevant case for
robotics. Depending on the representation language, literal values can be
naturally represented and processed (common case in logic programming language)
or not (storing numerical value, let alone matrices, in OWL is cumbersome and
not efficient due to the serialisation to XML).

The issue of the granularity of models can also be partially addressed by
splitting the knowledge representation into a geometric level (where all
numerical values are stored) and a purely symbolic level. The communication
between the two layers is however a complex question.

It must finally be noted that complex robotic systems are likely to use
different level of the knowledge depending on the task to achieve, and the
granularity of the knowledge should probably be considered as a dynamic
property.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Existing systems for knowledge representation in service robotics}
\label{sect|surveyed-systems}

Table \ref{table|surveyed-systems} lists the knowledge representation
systems that we have surveyed.

This section briefly presents each of them.

\begin{landscape}
\begin{table}\scriptsize
\begin{center}

\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}lp{2.4cm}p{3.4cm}p{1.5cm}}
\toprule
{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Project homepage} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
\midrule
ARMAR/Tapas & Formal & Holzapfel, Waibel \par (Karlsruhe TH) & & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Brimingham Univ., DFKI Saarbrücken) & & & Amodal proxies & \cite{Jacobsson2008} \\
GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & & \cite{Mavridis2006} \\
Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & \url{www.wrighteagle.org/en} & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU Munich) & \url{ias.in.tum.de/kb/wiki} & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & & NKRL & & \cite{Sabri2011} \\
%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & \url{incorl.hanyang.ac.kr/xe} & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
%ORO & KRS & Lemaignan, Alami \par (LAAS-CNRS) & \url{oro.openrobots.org} & {\sc Java} & OWL-DL ({\sc Jena}) & {\sc Pellet} & \cite{Lemaignan2010} \\
PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & \url{www.aass.oru.se/~peis} & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
%Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
% (based on {\sc KnowRob} & & (JSK) \\

\bottomrule

\end{tabular}
\end{center}

\caption{List of surveyed systems. Categories are \emph{Formal} for systems
that have a formal underlying knowledge representation, \emph{Ubiquitous} for
systems where knowledge is fully distributed, \emph{Language} for languages
used as KRS on robots or \emph{Structural} for KRS where knowledge is
represented as special data structures.}

\label{table|surveyed-systems}
\end{table}
\end{landscape}

%\fxfatal{Bielefeld -> could not find much\ldots}
%\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}

\subsection{ARMAR/Tapas}

{\sc Tapas} is the name of the knowledge representation system and dialogue
manager found on the ARMAR~III robot~\cite{Holzapfel2008} for the Karlsruhe
Institute of Technology.

Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
knowledge. The later is split into \emph{lexical knowledge}, \emph{semantic
knowledge} and a database of identified objects (with their properties). The
\emph{lexical knowledge} contains lexical and grammatical informations about
the objects. The \emph{semantic knowledge} is organised into an ontology
relying on \emph{typed feature structures} (TFS~\cite{Carpenter1992}, a
formalism originating from the computational linguistics community, and a
superset of first-order logic).

{\sc Tapas} has a strong focus on natural language grounding. It proceeds by
generating grammars from properties represented in the ontology to parse and
understand dialogue.

Another focus is put on handling unknown words and objects. {\sc Tapas}
provides routines to recognise unknown entities, and propose and interactive
and iterative verbal process to categorise (including adding new categories)
those new concepts.

\paragraph{Experiments} {\sc Tapas} has been used experimentally in a kitchen
environment where naive users had to ask the robot for an object and get
information about another object.

\subsection{CAST knowledge model}
\label{sect|cast}

CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
comprehensive toolkit aimed at building cognitive architectures for robots
through a set of interconnected \emph{SA} (\emph{subarchitectures}). The CAS
does not expose a central knowledge base as seen in other works. It represents
instead knowledge as unrooted \emph{proxies}. Those proxies are formally
defined in \cite{Jacobsson2008} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
\emph{proxies union} that form an equivalence class corresponding to one
entity.

A union of proxies forms a global amodal representation of an entity, that can
be explicitly shared and manipulated. Being not centralised, the knowledge
model can be qualified of \emph{ubiquitous}. Furthermore, knowledge source in
the CAS architecture is tightly bound to the on-line grounding process (be it
grounded in perception or in dialogue). While nothing seems to prevent it, no
{\it a priori} knowledge (including common-sense knowledge) is used.

Knowledge sharing is ensured by the event mechanism of CAST: modules can
monitor proxies for alteration by other modules. Jacobsson et al. mention how
this can apply to reinforcement learning: the vision module creates a proxy for
an orange object. This proxy get monitored by a learning module. In parallel,
the proxy is bound to an union by the natural language understanding module
that add new a feature like \emph{"this object is a fruit"}. The learning
module is called back, and can add this new information to its model.

In the presented implementation, the CAST knowledge model does no allow for
effectively representing actions or temporal information.\fxwarning{What about
reasoning? can they retrieve for example 'all proxies for colorful objects'?}

\paragraph{Knowledge Acquisition} Several techniques for knowledge acquisition
have been explored within the CAST framework. Cross-modal knowledge
fusion~\cite{Hawes2007a} is well studied, and the interaction with natural
language processing~\cite{Kruijff2010, Kruijff2010a} is a particular emphasise
of the project.

In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
the context of spatial representations with the robot \emph{Dora}.

\paragraph{Experiments} CAST has been used in several experiments, including
table-top manipulation (with a focus on language understanding) and more
recently on the Dora robot~\cite{Hawes2011} for indoor exploration.

\subsection{GSM}
\label{sect|gsm}

GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
representation system primarily built to ``facilitate cross-modal
interoperability'',  especially in the context of verbal interaction with a
robot.

GSM does not rely on any formal language but rather on a layered data structure
(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
relations between agents.  Each agent (any animate or inanimate object) is
attached to a physical model (made of \emph{body parts} that have properties
like their position, color, etc.) and a mental model (which is a recursively
embedded GSM, thus allowing a sort of theory of mind).

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{gsm.pdf}

    \caption{Simplified hierarchical structure of the Grounded Situation Model,
    based on~\cite{Mavridis2006}.}

    \label{fig|gsm}
\end{figure}

Properties are represented in three layers: a stochastic representation, close
to sensory percepts, a \emph{continuous single-valued} encoding of the
stochastic model, and a discrete, categorical model.

One notable feature of GSM is the \emph{bidirectionality} of the grounding
process: not only sensor percepts are abstracted into categories suitable for
human conversation, but human utterance (like ``There is a red ball in the
center of the table'') can also be turned into property descriptions. This
basically enable the knowledge representation system of the robot to
\emph{imagine} entities.

GSM also features several strategies for managing time and events.
\emph{Moments} are created by storing timestamped snap-shots of GSM, and
\emph{event classifiers} allow to define and detect events.

\paragraph{Experiments} GSM has mostly been tested on table-top manipulation
and interaction tasks (a ``conversational helping hand'' as stated by the
authors) implemented on a 7-DOF arm equipped with force feedback, cameras for blob
tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
an in-depth analysis of the performance of GSM by the mean of a standard
psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.

\subsection{Ke Jia Project}
\label{sect|kejia}

The Ke Jia project~\cite{Chen2010} integrates on a mobile platform a knowledge
representation language with natural language processing, task planing and
motion planing.

Knowledge representation relies on \emph{Action Language C}, itself based on
\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
are syntactically close to Prolog, are based on \emph{stable models} of logic
programs, and support non-monotonic reasoning. Default and non-monotonic
reasoning has been especially researched within the Ke Jia project for symbolic
task planing~\cite{Ji2011} and underspecified natural language processing.

Amongst other features, the natural language processing capabilities of the
system support acquisition of new logical rules at run-time.

\paragraph{Experiments} The Ke Jia robot has been demonstrated in several tasks
involving human-robot interaction with natural language. These tasks include a
task with multiple \emph{pick \& carry} that are globally optimised, naive
physics reasoning via taught rules or more complex scenarii with the robot
delivering drinks, taking into account changing and mutually exclusive
preferences of users.

\subsection{KnowRob}
\label{sect|knowrob}

{\sc KnowRob}~\cite{Tenorth2009a} is an integrated knowledge management system
developed at the Technical University of Munich. It is build as a set of
modules (figure~\ref{fig|knowrob} organised around a core reasoning system
written in Prolog. This core module interfaces through Java/Prolog or
C/Prolog APIs with external modules.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{knowrob.pdf}

    \caption{Overview of the {\sc KnowRob} framework, taken
    from~\cite{Tenorth2011}.}

    \label{fig|knowrob}
\end{figure}

Extension modules can plug into the system to provide specialised reasoning
capabilities or interfaces to external data sources, \eg to read object
detections from the vision system. These modules operate on the level of
instances (ABox).

\paragraph{Knowledge model} {\sc KnowRob} can load OWL ontologies, and the
\emph{KnowRob-Base} ontology is provided as a common-sense ontology, with a
focus on household and kitchen domains. {\sc KnowRob} also store and reason on
introspective knowledge through the \emph{Semantic Robot Description
Language}~\cite{Kunze2011} that allow to represent symbolically the
capabilities of the robot, and is used for planning.

\paragraph{Reasoning Techniques} Amongst the notable {\sc KnowRob} extensions,
{\sc ProbCog}~\cite{Jain2009} is an effort to provide probabilistic reasoning
based on bayesian networks, integration with naive physics reasoning has been
studied~\cite{Kunze2011a}, automatic parsing of Web resources in semi-natural
language has been also experimented with~\cite{Nyga2009}.

\paragraph{Grounding} {\sc KnowRob} offers a mechanism called \emph{computables} that allow to
evaluate certain predicates by calling external dedicated functions (for
instance, the valuation of a proposition like \stmt{object1 isOn object2} is
computed when required by calling a specific geometric reasoning module). In
combination with Prolog's lazy evaluation strategy, this supports a good
scalability.

Computables rely on various subsystem to evaluate. In particular, it relies on
the CoP framework~\cite{Klank2009} and \emph{semantic maps}~\cite{Blodow2011}
for the recognition of objects and environment.

%\paragraph{Integration with Decisional Layers} CRAM~\cite{Beetz2010}

\paragraph{Experiments} {\sc KnowRob} has been deployed on several scenarii at
the Technical University of Munich, on the PR2 robot and on a 2-arm custom
mobile manipulator in the scope of the ``assistive kitchen''~\cite{Beetz2008}
project. These experiments include retrieval and automated parsing of recipies
from the Web, retrieval and manipulation of various kitchen tools, cooperation
between two robots.

\subsection{NKRL}
\label{sect|nkrl}

\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
While this language is developed since a long time by Zarri~\cite{Zarri1997,
Zarri2008}, recent research direction include application to the robotic
field~\cite{Sabri2011}. NKRL is not {\it per-se} a knowledge representation
system, as it is primarily a language. However, it is used as the
representation and reasoning mechanism for robots by Sabri et al.

\paragraph{Knowledge representation} The NKRL language semantics are stored in
two ontologies: an ontology of concepts $\Omega$ and an ontology of events
$\Psi$. The ontology of events is made of action or situation templates.
Templates are a set of predicates (MOVE, PRODUCE, RECEIVE, EXPERIENCE, BEHAVE,
OWN and EXIST) associated to thematic roles. Grounding and reasoning with NKRL
is based on template matching.

\paragraph{Experiments} The main scenario of development for NKRL-based robots
is the Smart Home and monitoring of eldery people. Knowledge acquisition
partially rely on ambient intelligence (RFID, pressure sensors in the chairs,
etc.). The scenario is still being implementated.

\subsection{OUR-K and OMRKF}
\label{sect|omrkf}

The Ontology-based Unified Robot Knowledge~\cite{Lim2011} (OUR-K) framework,
successor of the Ontology-based Multi-layered Robot Knowledge
Framework~\cite{Suh2007} (OMRKF), is a knowledge representation system based on
five inter-related \emph{classes} of knowledge (figure~\ref{fig|omrkf}). It
proposes a layered approach to knowledge representation that allows to
integrate the grounding process to the knowledge representation process. OUR-K
knowledge model is implemented with a mix of Description Logics for the concept
hierarchies and Horn clauses.

\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{ourk.pdf}

    \caption{OUR-K organises knowledge into fives \emph{classes}, each composed
    of \emph{levels}. Figure based on~\cite{Lim2011}.}

    \label{fig|omrkf}
\end{figure}

Each level of knowledge is build as three stages of ontological realization: a
\emph{meta-concept} (the level itself, like ``temporal context'', ``behaviour''
or ``object feature''), a taxonomy of concepts inside this level (for instance
$cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
taxonomy ($cup1 : cup$).

\paragraph{Representation} The environment is represented in OUR-K in the
$spaces : Model$ knowledge level as a classical three layers mapping (metric,
topological and semantic maps). Objects (in $objects : Model$) are localised in
$spaces : Model$ through Voronoi nodes.

The knowledge class $Context$ proposes an explicit statement of spatial context
(mostly geometric relations between objects), temporal context and a more
general \emph{high-level} context, inferred from spatial and temporal contexts.

Finally, the $Activity$ knowledge class store compound actions in a HTN-like
structure, exploited at run-time by a planner.

\fxwarning{re-read the paper to improve the section...}

\paragraph{Experiments} Experiments conducted with OUR-K and OMRKF include
finding kitchen objects and reporting about their state to a human.  This
experiment also shows how OUR-K can deal with objects only partially matched by
their descriptor by introducing a $candidate()$ function.

\subsection{ORO}
\label{sect|oro}

\begin{figure}
    \centering
    \includegraphics[scale=0.36]{oro.pdf}

    \caption{Overview of the ORO server, taken from~\cite{Lemaignan2012}.}

    \label{fig|oro}
\end{figure}

\subsection{PEIS KR\&R}
\label{sect|peis-ecology}

{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that aim
to binds autonomous robotics with ambient intelligence (network of sensors).
\emph{PEIS} stands for \emph{Physically Embedded Intelligent System}: every
robots or intelligent device in the environment is abstracted as a PEIS.

Each PEIS physical component is running a \emph{PEIS Kernel} instance.
Communication between instance relies on a custom P2P communication protocol.

The PEIS architecture allows for adding new abilities through software
components sharing the common \emph{tuple space}.

We consider here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS
KR\&R}, that includes symbolic representation and reasoning.

\fxwarning{have a look to latest Daoutis paper!}

% More in details:
% - object identification based on viewpoint independent SIFT features
% - formalised anchoring system that explicitly match percieved attributes to predicates
% - Cyc predicates
% - ground 12 colors, based on a paper on color perception. Could be useful for us.
% - idem, they cite a paper on what spatial relations to compute
% - location of objects based on a previously provided semantic map (but not much on this semantic map)
% - two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
% - uses directly Cyc (ie, 250 000 common sense concepts\ldots), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
% Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
% - disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
% - template based natural language
% - experiment conducted in a "smart" indoor environmement + simple robot

\paragraph{Knowledge model} The PEIS Knowledge representation system relies on
the {\sc ResearchCyc} and {\sc CycL} language to represent knowledge. The {\sc CycL} language
allows to represent first order logic sentences and has extensions for modal logics and higher order logics.

\fxwarning{Is modal logics and higher order logics actually used in PEIS?}

As a system relying on {\sc CycL}, contexts can be expressed as
\emph{microtheories}: the truth or falsity of a set of statement depends of the
\emph{microtheory} in which these statements are evaluated.

\fxwarning{OWA/CWA?}

\begin{figure}
	\centering
	\includegraphics[width=0.9\columnwidth]{peis-architecture.pdf}
	\caption{The PEIS knowledge representation system, taken from~\cite{Daoutis2009}}
	\label{fig|peis-archi}
\end{figure}

The PEIS KR\&R system is deeply integrated to the general PEIS Ecology
\emph{smart} environment. Figure~\ref{fig|peis-archi} gives an overview of the
interactions between PEIS knowledge processing layers.

\paragraph{Knowledge Acquisition} The primary source for knowledge acquisition
is perception.  The PEIS ecosystem provides a SIFT-based object recogniser used
in conjunction with ceiling cameras for object localisation.  Other perceptual
modalities are available (like human tracking, ambient environment monitoring).

A template-based natural language parsing system may also be used to add new
assertions to the system.

The system can ask the human for help to disambiguate between concept names.

\paragraph{Anchoring} Daoutis et al. formalise the issue of anchoring as
finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
\Phi \times D(\Phi)$, where $\mathcal{P}$ is a set of predicate symbols, $\Phi$
a set of percept's attributes, and $D(\Phi)$ the domain of these attributes.

In the current implementation, object category (returned by the SIFT
classifier), color, location, spatial relations (both topological -- \emph{at},
\emph{near} -- and relative to the robot -- \emph{left}, \emph{behind}, etc.)
and visibility are the five classes of extracted attributes.

\paragraph{Integration in the robot architecture}
\label{sect|peis-integration}

The PEIS framework offers through the \emph{PEIS middleware} a practical way to
insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
module can be seamlessly integrated into the PEIS ecosystem.

\paragraph{Experiments} Experiments involving PEIS take place in a Smart Home
environment (\emph{PEIS Home}). The implemented case studies explore
dialogue-based interaction with the robot about known objects.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%% Underlying knowledge model table %%%%%%%%
\begin{table}
\begin{center}

\begin{tabular}{lp{4cm}}
\toprule
{\bf Project} & {\bf Common-sense \par knowledge source} \\
\midrule
{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
PEIS Ecology & {\sc ResearchCyc} \\
NKLR &  None \\
CAST Proxies &  None \\
GSM &  Predefined categories \\
OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
Ke Jia & None \\
ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\

\bottomrule

\end{tabular}
\end{center}
\caption{Underlying knowledge sources for each project}
\label{table|knowledge-sources}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Towards the next generation of Knowledge Representation Systems for Robotics}
\label{sect|conclusion}

This section tries to summary the various approaches we surveyed in the
previous sections and to draw some research perspectives. This section does not
exactly follow the list of compared features presented in
section~\ref{sect|features}. We have organized it along four axis: \emph{what
can be represented?}, \emph{How knowledge is created and grounded?}, \emph{What
can be done with the knowledge?} and \emph{How to use knowledge in the whole,
larger robot architecture?}.

The table~\ref{table|contribution-by-systems} sketches the landscape of current
research. The main fields of \emph{contribution} for the systems we have
surveyed are sorted along the different axis, dimensions of knowledge
representation that were identified in previous section.

\begin{landscape}
\begin{table}\tiny
\begin{center}
\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}|p{2cm}}
\toprule
\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}       & ORO \cite{Lemaignan2010}                      \\

\midrule

\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL + Horn clauses             & {\sc CycL}                    & DL (OWL)                                      \\
                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 &                               &                               & OWA                                           \\
                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         & + (\emph{candidate} entities) &                               &                                               \\
                                                                    & Meta-cognition &                           &                             &                             &                             & ++                                 & + (transformation rules)                        &                               &                               & ++ (reification, taxonomy walking)            \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                            
\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++                            &                               & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})\\
                                                               & Time representation &                           &                             & ++ (snapshots)              &                             &                                    & ++ (event oriented)                             & +                             &                               &                                               \\
                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            & +++                           &                               &                                               \\
                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          & ++                            & ++ (microtheories)            &                                               \\
                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 &                               &                               & ++ (Theory of Mind \cite{Warnier2012a})       \\
                                                                    & Self-knowledge &                           &                             &                             &                             & + (SRDF \cite{Kunze2011})          &                                                 &                               &                               &                                               \\
                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 &                               & +                             & +                                             \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                            
\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & +                             & ++                            & ++                                            \\
                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 &                               &                               & ++ (TBox alteration)                          \\
                                                                   & Lazy evaluation &                           &                             &                             &                             & +++ (Prolog, computables)          &                                                 &                               &                               &                                               \\
                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           &                                    &                                                 &                               &                               &                                               \\
                                                      & Presupposition accommodation &                           &                             & +++                         &                             &                                    &                                                 &                               &                               &                                               \\
                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             &                                    & ++                                              &                               &                               & + (explanation)                               \\
                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +                             & {\it++}                       & +++ ({\em HATP})                              \\
                                                           & Physics-based reasoning &                           &                             &                             & +                           & {\it+++} ({\em naive physics \cite{Kunze2011a}}) &                                   &                               &                               &                                               \\
%                                                                         & Learning & ++                        & ++ \cite{Jacobsson2008}     &                             &                             &                                    &                                                 &                               &                               &                                               \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                            
\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 &                               & ++ ({\em ambient intelligence})& ++                                           \\
                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 &                               & + (template-based)            & +++ ({\em Dialogs} \cite{Lemaignan2011a})     \\
                                                                     & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{Nyga2009}})&                             &                               &                               &                                               \\
                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & +++ (bottom-up)               & +++ \cite{Loutfi2008}         & ++ ({\em amodal model} \cite{Lemaignan2012})  \\
                                                              & Intrinsic Motivation &                           & ++ \cite{Hawes2011}         &                             &                             &                                    &                                                 &                               &                               &                                               \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                            
\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          & +                             & ++ (tuple space)              &                                               \\
                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++                            & ++ (tuple space)              & ++ (semantic events)                          \\
                                                          & Monitoring and debugging &                           &                             &                             &                             &                                    &                                                 &                               &                               & + ({\em remote visualisation})                \\
                                                           & Performances evaluation &                           & ++ \cite{Hawes2008}         & ++ (Token test)             & +                           & + \cite{Tenorth2011}               &                                                 &                               &                               &                                               \\

\bottomrule

%\end{tabularx}
\end{tabular}
\end{center}

\caption{Main domains of contribution of current KRS. Italics mean that the
feature is implemented as an external module. Main references are given in the
table header. When relevant, feature-specific publications have been provided.
An empty cell means that either the system has no specific focus on this domain
or we could not find relevant literature.}

\label{table|contribution-by-systems}
\end{table}
\end{landscape}

To comment this table, we propose select nine topic based on some of the
challenges that McCarthy and Roy have identified (there are listed at
section~\ref{sect|evaluation-literature}) on the road to natural interaction
and, to slightly paraphrase McCarthy, ``human-level robots''.

\subsection{Logic formalisms, continuous world}

McCarthy underlines the necessity of relying on mathematical logic, as the most
fruitful formalism for machine intelligence. This is largely the case. Almost
all the systems we have surveyed rely on logical formalisms, in several cases
as a mix of declarative languages (often based on Decription Logics) and
logical programming languages (Prolog, of course, but not only).

The two exceptions (the CAST knowledge model and GSM) are however also
interesting: CAST proposes a pervasive model of knowledge that is seducing from
the grounding perspective (but likely suboptimal for deliberative tasks), and
GSM encodes knowledge in an amodal (continuous, geometric) model while
preserving features that are usually specific of symbolic models (like theory
of mind or categorical knowledge).

The interleaving of discreet symbolic models with continuous geometric models
remains a major challenge, and techniques vary a lot: besides the CAST proxies
and the GSM amodal model, most systems directly extract physical attributes of
the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
adopt a top-down approach where local geometric models are set up on-demand to
compute symbolic properties like relative locations.

In our approach, ORO has tight (but unidirectional) links with SPARK, the
component in charge of geometric modeling and reasoning. Similarly to the GSM
approach, SPARK is an amodal model of the environment we can manipulate (3D
motion planing, for instance takes place in MHP, the twin brother of SPARK).

\subsection{Management of uncertainty and approximation}

McCarthy also mentions the necessity to deal with \emph{approximate concepts
and approximate theories} (that includes representing them and reasoning with
them).

The most notable attempt at modeling uncertainty down to the knowledge
representation is {\sc ProgCog}, an extension of {\sc KnowRob}. The theoretical
and practical performances (including decidability) remain however difficult to
overcome.

The challenge of representing and reasoning under uncertainty has not been
tackled in ORO server, at least not \emph{within} the current ORO server
knowledge model. Two reasons explain this absence: the decisional architecture
developed at LAAS for the robots has not consistent approach to uncertainty
representation and management, neither at the \emph{symbol production} level
(\ie SPARK or {\sc Dialogs}) or at the symbol consumption level (execution
control). Thus, no strong incentive pushed us in this direction. Then, the
current tools available from the Semantic Web domain to manipulate and reason
about knowledge do not provide mature support for uncertainty management. Some
effort do exist (like the {\sc Pronto} reasoner~\cite{Klinov2008}), but this
does not appear to be a major focus in the currently available tools.

\subsection{Non-monotonic reasoning}

McCarthy gives a particular importance to non-monotonic reasoning. Systems like
Ke Jia explicitly tackle this issue.

A monotonic system does not theoretically allow for knowledge retraction during
the reasoning, which is an important issue in the robotic context where the
world model is likely to be often altered.  However it is a practical issue
only if the reasoning process has to be \emph{continuous} during the whole
robot's activity lifespan, which is rarely the case. It is often possible to
stop the reasoner, alter the knowledge, and restart the inference process on a
new domain.

Non-monotonicity can also be partially dealt with with appropriate time
representation and reasoning (the reasoner then only takes into account
statements that are set as valid for a given moment).

Finally, probabilistic reasoning also implicitly leads to non-monotonic
reasoning, by relying on a continuous description of the state of the
world.

Non-monotonicity is however of broader significance to the knowledge
representation field, in particular for the representation of common-sense
knowledge where \emph{default} representation is currently sorely lacking.

It is also related to what McCarthy calls \emph{elaboration tolerance}: the
ability to extend \emph{on demand} the closed domain of interpretation for a
given assertion, to take into account new contextual knowledge.

\subsection{Modeling of contexts}

Both Roy and McCarthy mention the importance of formalizing and reasoning about
contexts. Many of the KRS we have surveyed mention at some point the context
modeling, but no consistent interpretation, let alone theory, of context
management has clearly emerged.

Several approaches for building contextualized knowledge have been presented in
this thesis: symbolic environment interpretation, perspective taking and
independent mental state for each agent, grounded natural language resolution,
self-awareness of its own activity. Much remains to be done for a robot to
actually identify its current context as well as contexts that may be referred
to.

We will further develop this important topic in the conclusion of the thesis.

\subsection{Reasoning about time, events and actions}

While representation of spatial roles (topology, placement, with or without
perspective-taking) is well studied and KRS usually integrate with spatial
perception and reasoning components, time representation and integration in
knowledge systems is less uniform.

The two main tasks that require time reasoning are task planing and sequence
recognition (in particular, action recognition). Systems that directly
integrate task planing (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
fluents in {\sc KnowRob}) to represent time.

Other approaches include storing snapshots of the knowledge state (like GSM),
that is used to move back to past mental states (of the robot or of another
agent).

Integration of these techniques with memory mechanisms (forgetting and
reinforcement learning) remains to explore.

Temporal modeling is currently absent of ORO, and symbolic and geometric
planning are accomplished outside of the main knowledge representation
layer. As a matter of fact, knowledge is mostly atemporal in the ORO server: the
set of triples stored in the server represents the beliefs of the robot
\emph{at the current time}. The robot represents knowledge about the \emph{here
and now}. Dealing with time reasoning is delegated to dedicated tools.

However, ORO does provide a simple implementation of a memory mechanism that
relies on time: when a statement is attached to a memory profile, the statement
is reified and the date of creation is stored. This allows to remove the
statement after a period of time that depends on the memory profile.

Statement reification to store timestamps of creation for all statements is
technically very easy to do in ORO server, but has not been actually enabled
because the performance hit (each reified statement produces four triples
instead of one) was not justified by any current use case of ORO server.

Regarding events, several systems (GSM, NKRL, ORO) have adopted an {\em
event-oriented} architecture where conditions (or templates) are used to
trigger decisional and execution processes.

A side note about how thematic roles and action models are initialized in ORO:
The current implementation relies on a small, predefined set of action verbs
that can be recognised from natural language
(section~\ref{processing_of_actions}).  This constraint does not come from the
resolution algorithm itself, but rather from the difficulty to automatically
extract the thematic roles associated to a verb. An interesting yet easy to
implement extension would consist in binding the dialogue processor with the
symbolic task planner. The task planner could dynamically provide the list of
actions that the robot can process, \ie actions for which the robot can produce
a plan. In certain case, it could also validate the understanding of a desire
by attempting to plan it.

Also, we could exploit on-line resources like
\textsc{VerbNet}~\cite{Kipper2008}, which provides a large machine-processable
lexicon of English verbs along with their thematic roles.

\subsection{Grounding in a multi-modal environment}

Symbol grounding remains a focus for most of the developers of cognitive
architectures for robots, and all the systems we have surveyed implement
grounding strategies (...it was an inclusion criteria). Most of the systems
adapt their grounding strategies to the sensing modality, so that grounding can
not be considered as a single, well delimited process.

Some systems (ORO, GSM) introduce an intermediate step in the grounding process
by the mean of an amodal model of the environment that aggregates the
perceptions (or suppositions) in a single place. This enables geometric
reasoning that takes into account all the perceptual modalities (typically
required for spatial perspective taking: we need to know where the humans are
looking at, and also where are the objects), and also improve the observability
of the system.

We note in addition that the development of what we have called ``synthetic
sensors'' is simplifying the grounding task. The most obvious example is the
human tracking system provided by the low-cost Kinect device, and now used
pervasively in the robotics lab. Not only this system segments and computes the
pose of humans that enter its field of view, but it also tracks them (including
in case they are partially or completely occluded). This kind of high level
sensing device does not completely remove the need of grounding one symbolic
instance of a human with the physical human, but the task is ways simpler than
it used to be.

Many systems also tackle the difficult question of natural language processing.
The CAST middleware, in particular, has been used in the European CoSy and CogX
projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
the huge interest for the grounding of verbal interaction within the cognitive
robotics community, and we are likely to reach important milestones in this
field in the coming years.

Merging more modalities (especially, back-channel communication, deictic
gestures and social gazes) also sparks a lot of attention, and is becoming more
and more present as symbolic knowledge available to the control layers.

\subsection{Common-sense}

McCarthy actually starts his list by affirming that intelligent systems must be
able to \emph{``operate successfully in the common sense informatic
situation''}. This question of the \emph{common-sense} is probably one of the
toughest because what \emph{common-sense} mean is not very clear at first place
(since, by definition, common-sense is, well, common-sense...). We feel,
however, that it is related to a diffuse cultural background, and one may even
claim that, for a system to acquire common-sense reasoning is equivalent to
solve the strong AI challenge (this is at least more or less McCarthy's opinion).

Our knowledge representation systems have a pragmatic and potentially very
powerful approach to common-sense: reuse knowledge stored on the Web. While not
so many of the systems we have surveyed directly tackle this question ({\sc
KnowRob} is the only one explicitly working on common-sense reasoning, through
physics-based reasoning, integration with Web databases or parsing of
semi-structured Web documents), several rely however on Web standards (OWL,
OpenCyc) to represent their knowledge. With the development of initiatives like
OpenMind that encode with the same standards common-sense facts and rules,
there is a strong potential for our robot to gain common-sense knowledge from
well-structured online resources in the coming years.

%%%%%%%%%% Underlying knowledge model table %%%%%%%%
\begin{table}
\begin{center}

\begin{tabular}{ll}
\toprule
{\bf Project} & {\bf Common-sense knowledge source} \\
\midrule
ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\
CAST Proxies &  None \\
GSM &  Predefined categories \\
Ke Jia & None \\
{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology, physics simulation \\
NKLR &  None \\
OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
PEIS Ecology & {\sc ResearchCyc} \\

\bottomrule

\end{tabular}
\end{center}
\caption{Underlying common-sense knowledge sources for each project.}
\label{table|knowledge-sources}
\end{table}

Table~\ref{table|knowledge-sources} lists for the systems we have surveyed the
current sources of common-sense knowledge.

\subsection{Learning, representation of experience, introspection}

Learning has not been addressed in this work. This is a complex, still emerging
field, whose exact meaning and perimeter varies a lot. It was difficult to
synthesise in the KRS taxonomy, and we decided to omit it. We did not either
really tackled this question in ORO, except in term of knowledge structure
alteration based on verbal interaction (this is presented in one case-study, in
the next section).

Learning at the level of the knowledge representation system has been explored
in conjunction with visual perception systems (mentioned for instance, in the
CAST project, in~\cite{Jakobsson2008}). {\sc KnowRob}, by filling its pool of
facts from informations automatically extracted from the Web~\ref{Nyga2009},
can also be considered as a \emph{learning} system (it autonomously acquire
knowledge). We lack however a formal study of learning strategies and
techniques at the knowledge representation level.

Representation and matching of past experience is a related topic. This ability
is a key step for general action recognition, and is of particular importance
for the robot to assess the state of the interaction with the human.

While we already mentioned that several systems are able to reason on past
states, we are not aware of existing implementation of algorithms to reflect on
past experiences.

This is itself related to introspection and meta-cognition: the shift towards
explicit knowledge representation exemplified in the nine systems we have
presented, has a major impact on the meta-cognitive capabilities of our robots.
They can exhibit, manipulate and reason on their internal belief state (what we
will call in the conclusion the \emph{cognitive observability}). We have still
to discover all the possibilities that are open by this important cognitive
ability.

\subsection{Perspective-awareness}

Finally, Roy mentions \emph{the ability to take into account the human
perspective}: this cognitive ability, that relates to the representation of
different modalities, is present in ORO and GSM. ORO explicits the perspectives
in different symbolic mental states while GSM recursively stores models for
each agent.

We have already explained how perspective awareness enables advanced cognitive
capabilities like a theory of mind, and this is probably the main contribution
of this thesis regarding techniques for knowledge representation.

% \subsection{Justification and resolution of cognitive conflicts}
% 
% Logical inconsistencies in one ORO's model (or, from a cognitive point of view,
% cognitive conflicts) could be better dealt with: while (partially) provided by
% the Pellet reasoner, and output to the developers for debugging purposes, the
% logical justification of the inconsistency is not really exploited.

% \subsection{Performances analysis}
% 
% \subsubsection{Relaxation time}
% 
% The relaxation time (as defined at section~\ref{sect|performances}) of ORO for
% the average workloads we have encountered during the experiments is in the
% range of 100ms.  The temporal resolution that follows for the symbolic layer,
% while acceptable in most cases, is not satisfactory. In particular, multi-modal
% interaction requires fine grained temporal alignement between
% modalities~\cite{Li2012} which are not possible with the current temporal
% performances of ORO.
% 
% \subsubsection{Cognitive evaluation}
% 
% The performance evaluation of current tools is still mostly system- and
% scenario-specific (``did my system do the job for my task?''), and amongst the
% surveyed KRS, only few systems proposes an evaluation of their performances
% that could be compared to other system.
% 
% 
% \fxwarning{Mention reasoning and expressiveness, cf section oro-server}
%

\subsection{Towards a new design}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{ieeetr}
\bibliography{biblio}

\end{document}