### 2014-04-21 krs-survey.tex
reworked text, added references

\documentclass[journal]{IEEEtran}
\usepackage[utf8]{inputenc}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subfigure}

% typology diagrams
\usepackage{tikz}
\usetikzlibrary{shapes}

%tables
\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{multirow}
\usepackage{varwidth}

\newcommand{\turn}[3][10em]{% \turn[<width>]{<angle>}{<stuff>}
  \rlap{\rotatebox{#2}{\begin{varwidth}[t]{#1}\bfseries#3\end{varwidth}}}%
  }
\usepackage{pdflscape} %% Used for very big table

\usepackage{paralist}

\usepackage[draft]{fixme}
\fxsetup{footnote}
\fxsetup{nomargin}

\graphicspath{{figs/}}

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf~}}}
\newcommand{\eg}{{\textit{e.g.\ }}}

\newcommand{\taxon}[2]{%
    \textbf{[#1] \emph{#2}}
}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:} #1}}

% Macros related to concepts/statements typesetting
\newcommand{\concept}[1]{{\footnotesize \texttt{#1}}}
\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
\def\stmttt#1 #2 #3\relax{{\tt#1} {\bf{\tt #2}} {\tt #3}}
\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
\def\setstmttt#1,#2\relax{\rawstmt{#1}, \rawstmt{#2}}

\title{Knowledge Representation for Service Robots: A Survey}
\author{Séverin Lemaignan, Moritz Tenorth, Rachid Alami, Michael Beetz}

\begin{document}

\maketitle

\begin{abstract}

The aim of this article is two-fold: We first propose a typology of
knowledge representation systems for service robotics that covers intrinsic
features, such as expressiveness or reasoning techniques, as well as extrinsic
features, such as knowledge acquisition and integration into pre-existent 
robotic architectures. We base the typology on an initial typical scenario of 
service robotics.

We then survey current knowledge representation systems in the community to
identify which knowledge-bound cognitive abilities are today well understood
and implemented, and which ones require more research efforts.

We hope this article may bring a clearer picture of the challenges of knowledge
representation and manipulation to the robotic research community.

\end{abstract}

\fxnote{Support material: \emph{What is a knowledge representation} by Davis,
Shrobe and Szolovits,
\url{http://groups.csail.mit.edu/medg/ftp/psz/k-rep.html}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sect|intro}

\todo{MT: For a T-RO paper, we should maybe start a bit differently and 
      emphasize the relevance for robotics and the timeliness of such an 
      overview more?}

\todo{Other ideas:}

\begin{itemize}
  \item Check if/which systems are available as open-source software?
  \item KnowRob: add more recent developments like reasoning about 3D CAD models
  \item 
\end{itemize}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Knowledge as the Root of Cognitive Abilities}
\label{sect|cognitive-abilities}

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Foundations of Artificial 
Intelligence}~\cite{Levesque2008}, Levesque reminds the reader of the manifesto 
they wrote together in 1998:

\begin{quotation}

    ``Central to this effort is to develop an understanding of the relationship
    between the knowledge, the perception, and the action of [\ldots] a robot. The
    sorts of questions we want to be able to answer are

    \begin{itemize}

        \item to execute a program, what information does a robot need to have
        at the outset versus the information that it can acquire \emph{en route}
        by perceptual means?

        \item what does the robot need to know about its environment versus what
        need only be known by the designer?

        \item when should a robot use perception to find out if something is
        true as opposed to reasoning about what it knows was true in the past?

        \item when should the inner workings of an action be available to the
        robot for reasoning and when should the action be considered primitive
        or atomic?

    \end{itemize}
    \noindent
    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.''

\end{quotation}

To be answered, those questions have one prerequisite in common: The robot must
be able to \emph{explicitly} manipulate knowledge, and hence, needs a way to
represent the knowledge.
% 
This survey focuses on this knowledge representation issue: We first aim at
establishing a comprehensive typology of representational needs for
robotics in the specific context of service robotics, and then at painting the
current landscape of approaches in the research community in order to summarize 
the major trends and to identify possible shortcomings.

We have classified the features of knowledge representation systems (KRS) into
six broad categories: \emph{What can be represented?} or the
\emph{expressiveness} question ; \emph{How are things represented?} ; \emph{Which
reasoning techniques are offered?} ; \emph{How is knowledge acquired and
anchored?} ; \emph{How is the knowledge made available and used by other
components of the robot?} and \emph{Which actual knowledge is stored?} or the
\emph{knowledge instantiation} question.

We have identified over ten projects that are either explicitly advertised as
knowledge representation systems or involve explicit knowledge manipulation
within a robot. By knowledge manipulation, we mean symbolic representation of
assertions, be it static statements on the world or spatio-temporal events, and
reasoning on this representation. Prominent features of each of these systems
illustrate how the various features of a knowledge representation system are
currently implemented.

We have also tried, for each project, to make clear \begin{inparaenum} \item
how information from perception or interaction is turned into knowledge,
\item how, in return, symbolic concept are \emph{anchored} into the robot's
sensori-motor space, \item how the knowledge representation system integrates
with the decisional layers of the robot, and how this can improve or ease
robotic control.\end{inparaenum}
\fxfatal{Do we really do it?}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\paragraph{What do we call ``knowledge''?}
\label{sect|on-knowledge}

Since we will discuss the concept of knowledge in the context of robotics in 
depth over the course of the the following pages, it is useful to make our 
terminology explicit first.
% 
Be it in philosophy, cognitive sciences or computer sciences, reaching an
agreement on a definition of ``knowledge'' seems difficult.

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, and \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language and semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the article, give a second look to
this distinction.

In our robotic context, we define knowledge as a narrower concept, while
keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
interrelated logical facts that are meaningful to the robot executive
controller}. By \emph{meaningful} we mean that they can possibly be interpreted 
to lead to a purposeful action. We will see that our main challenge while
designing a cognitive architecture is furthermore to make this knowledge as
\emph{explicit} as possible.

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question (with D,I,K 
and W standing for ``data'', ``information'', ``knowledge'' and ``wisdom'', 
respectively). In this article, we will associate ``data'' with low-level 
material like raw sensor output, and ``information'' with uncontextualised 
symbolic facts.
To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, coords)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot, would be proper knowledge.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection*{Survey Inclusion Criteria}
\label{sect|inclusion-criteria}

Every robotic system includes, explicitly or not, some kinds of knowledge 
representation systems. They may range from a simple state vector to an 
explicit symbolic knowledge base. This survey focuses on the right end of 
this spectrum: symbolic systems, suited for abstract reasoning.
% 
Besides, we have decided to restrain the set of systems to those actually
implemented on robots, and used in semantically rich environments (\ie dynamic,
partially unknown environments with a large range of different entities which
may have interactions). A typical scenario that would involve such robots is
a service robot in a human-friendly environment like a kitchen.
% 
We have further limited ourselves to systems that
\begin{inparaenum} 
  \item  run on a \emph{service robot} (that is, robots that interact with 
  objects in a semantically rich environment primarily designed for humans),
  \item  ground the knowledge in the physical world (physically embedded
  systems able to assess their environment),
  \item  are able to merge different knowledge modalities,
  \item  are able of on-line, dynamic knowledge acquisition and reasoning 
  (\ie not simple static databases).
\end{inparaenum}

\todo{MT: I think we should rethink this decision. DyKnow is a knowledge 
      processing system running on physical robots, and there's also some
      work by Lakemeyer etc in Aachen on using Golog-like things on household 
      robots}
These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.
% 
We have also chosen not to include the GOLOG language and its
derivatives~\cite{levesque1997golog, Ferrein2008, Gspandl2011} in this survey. While
several implementations on robots, including service robots, do exist, the
focus of this language is on representation and reasoning about actions and
situations, and the link with symbolic, abstract knowledge is not explicit.

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
GLAIR~\cite{Shapiro2009} or {\sc ACT-R} do have declarative knowledge
modules~\cite{Derbinsky2010} and have recently been used on service robots (see
 for instance {\sc ACT-R/E}~\cite{Kennedy2009}), they are also absent from this
survey because we did not find many references in the literature on knowledge
manipulation and representation applied to real-world robotic scenarii for
these architectures.
% 
A comprehensive reference on (bio-inspired) cognitive architectures is
available from the BICA Society~\cite{BicaCogArch2011}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Previous work}
\label{sect|evaluation-literature}

The typology we propose is partly based on a comprehensive synthesis of
classifications and analysis found in the literature. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation in the
context of service robotics.

In their chapter on Cognitive Robotics, Levesque and Lakemeyer~\cite{Levesque2008} 
present several characteristics of knowledge representation systems
for robots, stressing the need of representing the \emph{dynamics of the
world}.  Sensing is included in the knowledge representation via
\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
distinct parallel mental models; action representation and reasoning about
tasks is discussed in the context of \emph{situation calculus}; \emph{open
world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
robot programming and knowledge representation can be related. We integrate
most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define a
\emph{knowledge processing middleware} as a system supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context-dependent processing at many different levels of abstraction''. They
identify six characteristics: The system must be able to \emph{merge
information} from different, possibly distributed sources, it should support
quantitative as well as qualitative processing of information, it should offer
\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
(which requires what we here call \emph{non-monotonicity}), and finally support
\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
specification of the processing functionalities'').

Several surveys compare general cognitive architectures \cite{Langley2006,
Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
distinguish nine capabilities: recognition and categorisation, decision making,
perception and situation assessment, prediction and monitoring, planning,
reasoning, execution control, interaction and
learning/remembering/introspection. They also separately identify four
\emph{properties} of a cognitive architecture that categorize how knowledge is
handled by the architecture: representation of knowledge, organization of
knowledge, utilization of knowledge, and acquisition and refinement of
knowledge. This categorization had a notable influence on our typology, and
many of these categories are also present in our proposal.

Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, with a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology}
(how the structure of the world is captured in a representation and shared),
the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.

Chong et al.~\cite{Chong2009} also provide a recent review of the main cognitive
architectures, with a focus on eight functions: perception, memory, goals
management, problem solving capabilities, planning, reasoning, learning and
links to neurobiology.

At an even broader scope, several authors from fields that are connected to
robotics have previously listed desirable features of artificial systems aiming
at rich cognitive abilities. For instance McCarthy recently listed the 
challenges he identified on the road to a \emph{human-level AI}~\cite{McCarthy2007}.

\begin{itemize}

	\item The ability to \emph{"operate successfully in the common sense
	informatic situation"},

	\item The necessity of relying on mathematical logic, as the most fruitful
	formalism for machine intelligence,

	\item The ability to deal with \emph{approximate concepts and approximate
	theories} (that would include representing them, and reasoning with them),

	\item Non-monotonic reasoning,

	\item What McCarthy calls \emph{Elaboration Tolerance}: the ability to
	extend \emph{on demand} the closed domain of interpretation for a
	given assertion,

	\item The ability to formalize and reason about contexts,

	\item Reasoning about events, and in particular, actions,

	\item The capacity of introspection,

	\item and finally, he points at the issue of giving computer the right
	heuristics for decision making.

\end{itemize}

Coming from the perspective of natural language processing in situated context,
Roy and Reiter summarize in~\cite{Roy2005} what they see as the main challenges
to be tackled by knowledge representation systems: \emph{Cross-modal
representation} systems, association of words with perceptual and action
categories (\emph{grounding}), modeling of \emph{context}, definition of the
right \emph{granularity} of models, integration of \emph{temporal modeling and
planning}, ability to \emph{match past (learned) experiences} with the current
interaction and ability to take into account the \emph{human perspective}.
% 
Knowledge representation systems in robotics are directly affected by these
points, and we indeed integrate them in our typology, in slightly reformulated
ways.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Article overview}
\label{sect|overview}

The paper is organized in three main sections.
% 
Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
representation in robotics. It introduces briefly a range of different
approaches that do not focus on a specific field of robotics, and instead
provide insights on the wealth of knowledge representation techniques available
to researchers.
% 
Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
systems deployed on \emph{service robots}: Section~\ref{sect|features}
introduces a extensive typology of requirements and features of knowledge
representation systems, that we use as foundation to compare and evaluate the
systems we examine, and section~\ref{sect|surveyed-systems} then presents each
of the eight systems, and underlines their specificities.
% 
We finally summarize in section~\ref{sect|conclusion} the current approaches and
we attempt to identify several research directions that are not sufficiently
addressed by the current state of the art.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section{Knowledge Representation in Robotics: an Overview}
\label{sect|krsoverview}

Representations for a robot's knowledge have been investigated for decades, 
starting with the seminal work on the Shakey~\cite{nilsson84shakey} robot --
which already had an internal world model based on predicate logics. Shakey
however lived in a very simple artificial world consisting of a few blocks
that could rather easily be described in logical form. Scaling towards more
complex settings required research in both AI and robotics, and in the 
following decades, these two communities operated without much interaction.
Today's robots are now able to operate in much more complex environments 
compared to Shakey's blocks world, have perception techniques that can handle 
the complexity of real scenes, manipulation capabilities that allow them to 
interact with objects, and massively more computational resources. On the one 
hand, this makes robots an interesting application for AI research again. On
the other hand, roboticists strive for more flexibility, adaptability and in 
general more semantic representation and therefore consider AI methods again. 
Besides the large-scale knowledge representation systems that will be discussed 
in the following sections, there have been many approaches to using knowledge
in parts of the robotic system.

One field that has seen much progress in the past years is the creation and 
representation of environment models, often termed ``semantic maps''. Advances 
in object recognition allowed to identify single objects in a map, which 
suggests to also store higher-level semantic information about them. The 
term ``semantic'' is however used for a range of things -- from a mere 
classification into different parts and object types~\cite{Rusu08RAS2}or the 
co-occurrence of objects in scenes~\cite{vasudevan08} up to environment 
representations in description logic~\cite{zender2008conceptual}, statistical 
relational environment models~\cite{limketkai05relational} and an embedding 
of spatial information into an encyclopedic and common-sense knowledge 
base~\cite{tenorth10envmodel}. The use of semantic maps as knowledge source 
for task planning has been investigated by~\cite{galindo08taskplanning}. 
% 
Another area of research that often includes some knowledge representation
aspects is natural-language interaction. For understanding spoken dialog
about things in its environment\cite{Mavridis2006} and for following spoken
directions \cite{kollar10natural,matuszek12commands,duvallet13imitation}, a 
robot needs to be able to ground the symbols (i.e. words) in percepts and 
actions. The natural-language text may be the result of direct interaction 
with a human, but can also be obtained from resources on the Web, for example
for mining object locality knowledge~\cite{zhou12webmining} or extracting 
task instructions from Websites~\cite{tenorth10webinstructions}.
% 
Recently, learning from experiences and ``life-long learning'' have
become popular topics in robotics. The goal is to record rich episodic 
memories during task execution~\cite{niemueller2012iros}, to reason about 
them later to extract more generic rules~\cite{winkler13memory} and to learn
generalizes task models from the collected experiences~\cite{rockel13ontology}.

\todo{add something on the IEEE standard draft for a robotics and automation ontology}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

% We propose to organize requirements from an imaginary cooking scenario where
% numerous desirable features for a knowledge representation system for robotics
% have been identified without any particular order.
% 
This section proposes a formal typology and a nomenclature of analysis 
dimensions for a knowledge representation system. For each feature, we provide 
a short definition along with links to relevant literature. As mentioned 
earlier, we propose six main categories that cover both basic and practical 
aspects: The formal \taxon{A}{Expressiveness} of the system, the breadth and 
nature of the \taxon{B}{Representations} allowed by the system, the
\taxon{C}{Reasoning} capabilities, its methods for knowledge 
\taxon{D}{Acquisition}, the \taxon{E}{Integration} of the system with other 
components of the robot software architecture, and finally what we call
the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
base.

The next section details only briefly the exact meaning of these categories
and, accordingly, introduces sub-categories. We kindly refer the interested 
reader to~\cite{lemaignan2012symbolic} for an in-depth discussion of each of 
these dimensions of knowledge representation systems as well as related 
literature.
% 
Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
the main contributions of each of the surveyed systems with respect to these
dimensions. \todo{MT: Can we list this taxonomy of capabilities as one of the 
contributions of this article?}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Representational expressiveness}
\label{sect|expressiveness}

This first axis of analysis is the intrinsic expressive power. It answers the
question of what can possibly be represented. In case it explicitly exists, the
\emph{language} of representation obviously plays an important role here.
% 
We subdivide the category \taxon{A}{Expressiveness} into five subcategories:
\taxon{A.1}{Logical formalism} classifies the underlying logical formalism of the
KRS (Horn clauses like for Prolog-based systems, Description
Logics~\cite{Baader2008}, Bayesian logics, etc.). \taxon{A.2}{Expressive Power}
quantifies the design choices made with respect to the actual expressive power
of the logical formalism: tradeoffs have to be found to allow both for high
expressiveness and tractable and decidable representations, and KRS authors
often add constraints on their representation logics to this end.
\taxon{A.3}{OWA/CWA} reflects the design choice between the \emph{open world}
(OWA) versus the \emph{closed world} (CWA) assumptions, which refers to the
possibility of representing \emph{unknown} facts. \taxon{A.4}{Uncertainty}
relates to the ability of the system to represent both the \emph{intrinsic}
uncertainty of given facts, and the \emph{extrinsic} uncertainty due to
imperfect perception by the robot. \taxon{A.4}{} relates only to the possibility
of \emph{representation}. Reasoning on uncertain facts is discussed in
\taxon{C.4}{}. The last feature of a KRS related to representation is
\taxon{A.5}{Meta-cognition} that conveys the capability for a system to
represent and exhibit it own internal structures and representation mechanisms.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Representation techniques}
\label{sect|higher-level-domain-representation}

This category focuses on questions that involve representational challenges 
(time, space, context) or require specific cognitive capabilities (theory of 
mind, introspection, memory).
% 
The dimension \taxon{B.1}{Roles}, with its sub-categories \taxon{B.1.1}{Space},
\taxon{B.1.2}{Time} and \taxon{B.1.3}{Actions}, deals with the approaches that
the KRS suggests to represent these three important \emph{roles} in robotics
(spatial relations, time and representation of actions). To give a few examples
(in no way exhaustive), \emph{spatial relations} cover techniques like
topological maps, allocentric and egocentric frame projections. \emph{Time
representation} includes Allen's interval algebra~\cite{Allen1984}, Ghallab's
chronicles~\cite{Ghallab1996} or the concept of \emph{fluent}. \emph{Actions}
are concerned with representation of \emph{pre- and post-conditions},
\emph{thematic roles} and \emph{plans}.

\taxon{B.2}{Context} describes if and how a given KRS represents the
\emph{interpretation frame} of the situation and its own knowledge. This is
often related to the initial \emph{common-sense knowledge} available to the
robot. \taxon{B.3}{Modality} refers to the ability of the KRS to represent
possible alternate worlds or belief states. A typical application of a
\emph{modal} knowledge model is the implementation of a \emph{theory of mind},
\ie an independent belief model for other agents the robot interact with.
\taxon{B.4}{Introspection} covers both \emph{self-knowledge} (what do I know
about myself?) and a weaker version focused on the explicit representation and
reasoning on the \emph{robot's own skills} (can I execute a given task?).
Finally, \taxon{B.5}{Memory} refers to the implementation of
different kinds of (bio-inspired) memory models like \emph{declarative},
\emph{procedural} or \emph{episodic}, and the attached processes
(\emph{forgetting}, \emph{selectively remembering}, etc.).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Reasoning techniques}
\label{sect|reasoning}

As one can expect, many reasoning techniques born in artificial intelligence
have found their place in robotics over the years. We propose to consider ten
subcategories of \taxon{C}{Reasoning} that reflect features we consider 
important with respect to the general requirements of robotics (like tractable 
decision making in pseudo-real time, dealing with partially unknown or uncertain 
domains, or reasoning related to interaction with the physical world). Again, 
the descriptions of these ten subcategories hereafter are very superficial. 
Readers are referred to~\cite{lemaignan2012symbolic} for details.

\taxon{C.1}{Standard reasoning} covers conventional reasoning techniques based
on logical inferences (for instance, \emph{forward/backward chaining} or
\emph{semantic tableaux}). The main problems that these techniques address are
\emph{concept satisfiability}, \emph{consistency checking} and \emph{instance
checking}. \taxon{C.2}{Instantiation and structural alteration} describe how
dynamic a system is with respect to its internal knowledge structure (admitting
the modification of the knowledge structure, often called \emph{TBox}, may
for instance influence the capability of a system to \emph{learn}).
\taxon{C.3}{Lazy evaluation} describes the ability of a KRS to delay active
knowledge acquisition or reasoning operations until the value is actually
needed, and may have a strong impact on the overall scalability of the system.
\taxon{C.4}{Uncertainty} groups all the features related to reasoning over
uncertain knowledge (like probabilistic inference).
\taxon{C.5}{Non-monotonicity} discusses how systems deal with non-monotonic
inferences, \ie exceptions to a rule or inferences that lead to the retraction
of previously inferred facts. Techniques to deal with non-monotonic reasoning
include for instance \emph{Default Logics}.  \taxon{C.6}{Presupposition
accommodation} is the ability of a system to automatically create a context
allowing to make sense of a proposition. This is tightly linked to the 
capability of dealing with \emph{under-specification}. \taxon{C.7}{Prediction 
and explanation} (with subcategories \taxon{C.7.1}{Projection}, 
\taxon{C.7.2}{Legality}, \taxon{C.7.3}{Diagnosis} and \taxon{C.7.4}{Explanation}) 
relates to the capabilities of a system to predict
and project the outcomes of a sequence of actions onto the logical system, and
conversely, to justify and explain the sequence of entailments that lead to a
conclusion. Related to this category, \taxon{C.8}{Planning} discusses to which
extend a KRS can be used for \emph{symbolic task planning}, \ie the ability of
a robot to select a sequence of actions in order to reach a given final state.
\taxon{C.9}{Physics-based reasoning} describes the capability of a symbolic
knowledge representation system to rely on physical reasoning (which may include
geometry, dynamics, collision checking, etc.) to derive symbolic inferences.
Finally, \taxon{C.10}{Learning} groups the various capabilities that lead to
learning behaviors.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Knowledge acquisition}

In the survey inclusion criteria, we have chosen to only consider robotic
systems that acquire knowledge by themselves at runtime. The category
\taxon{D}{Acquisition} reviews the different approaches to knowledge acquisition
and semantics extraction. \taxon{D.1}{Acquisition and fusion} examines the
techniques to first acquire new logical statements and then \emph{align} them to
the corpus already available to the robot. Depending on the source of knowledge,
we distinguish three further subcategories: \taxon{D.1.1}{Sensing} covers the
techniques that translate perceptions into symbolic knowledge,
\taxon{D.1.2}{Interaction} focuses on knowledge specifically stemming from
interaction (typically, language processing), and \taxon{D.1.3}{Linked
Resources} relates to the connections and alignment issues with external
(often Web-based) knowledge bases.
% 
\taxon{D.2}{Grounding} discusses the KRS' capabilities with regard to attaching a
practical meaning to the symbol it manipulates. In the context of robotics, it
mainly consists of building and maintaining a bi-directional link between
sub-symbolic representations (sensor data, low-level actuation) and symbolic
representations.
% 
Finally, \taxon{D.3}{Intrinsic Motivation} describes the possible
\emph{internal} mechanisms by which the robot seeks by itself to acquire more
knowledge.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Practical integration in robotic architectures}
\label{sect|integration-robot}

Knowledge representation systems have limited usefulness for robotics if they are
considered in isolation, and it seems important to also evaluate KRS from the
perspective of their integration into larger control architectures that include
perception routines, decision-making processes and actuation control.
\emph{Practical} integration also suggests to take into account real-world
constraints, like performances and monitoring tools that come along with the KRS.

\taxon{E}{Integration} is hence subdivided: The integration of the KRS with the
\taxon{E.1}{Sensori-motor} layers may adopt various paradigms (from passive
\emph{semantic blackboards} to active KRS that directly query the sub-symbolic
layers. \taxon{E.2}{Executive layers} examines conversely how the KRS is used
from the perspective of the robot controller. Here also, many different
approaches have been investigated, from classical client-server architectures, 
to domain-specific language extensions. \taxon{E.3}{Monitoring} details the 
support tools provided by the system for the developers to monitor and 
introspect their system. This is especially important for knowledge representation 
systems because they often lie at the heart of the architecture and act as hubs 
for many other components. \taxon{E.4}{Performance}, finally, evaluates the
performance of the KRS -- the \taxon{E.4.1}{Raw performance} in terms of 
scalability and classification speed, the \taxon{E.4.2}{Cognitive performance} 
in terms of cognitive abilities provided by the system. Note that the performance 
of a KRS is notoriously difficult to assess since, on the one hand, cognitive 
architectures for robotics are often build upon tightly interleaved modules 
which makes it difficult to test a KRS in isolation, and on the other hand, 
because standard benchmarks and scales for cognitive capabilities are still 
lacking.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Knowledge instantiation}

The last category of our taxonomy looks at the actual \emph{content} of the
knowledge base: the knowledge instantiation. Here, \emph{instantiation} does
not only refer to the instantiation of the knowledge structure (what we have
called the ABox), but also includes the knowledge structure itself (the TBox).
% 
The \taxon{F.1}{Design strategy} of a KRS can be \emph{top-down} (from an
initial symbolic model down to the percepts), \emph{bottom-up}, or a mix of
both. The bottom-up and top-down strategies reflect how the knowledge structure
itself is constructed: either from generic categories, typically extracted from
standard upper-ontologies, or by successive classification and refinement of
percepts.
% 
\taxon{F.2}{Common-sense and alignment} discusses if and how the KRS reuses
existing pools of knowledge like \emph{upper ontologies}, and how the question
of knowledge \emph{alignment} is handled. \taxon{F.3}{Metrics and quality
criteria} includes qualitative and quantitative metrics that provide insight 
into the size, complexity and effectiveness of knowledge bases available to the 
robot. These metrics include for instance the number of terms and assertions, 
but also the type of predicates that are used, or the computed Description Logics
expressiveness (if applicable). The qualitative evaluation of ontologies is a
research field in its own right that has introduced criteria like \emph{accuracy,
adaptability, clarity, completeness, computational efficiency, conciseness, and
consistency}. The last category is the \taxon{F.4}{Granularity}, which accounts 
for the level of details considered in the knowledge representation system 
which can go down to storing SIFT features~\cite{Suh2007} directly in the KRS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Existing systems for knowledge representation in service robotics}
\label{sect|surveyed-systems}

Table \ref{table|surveyed-systems} lists the knowledge representation
systems that we have surveyed. This section briefly presents each of them.

\begin{table*}\scriptsize
\begin{center}

\begin{tabular}{p{2.2cm}p{1.6cm}p{4cm}p{2.4cm}p{3.4cm}p{1.5cm}}
\toprule
{\bf Project} & {\bf Category} & {\bf Authors (Institution)} & {\bf Programming language} & {\bf Knowledge model/Logical Formalism} & Main reference \\
\midrule
ARMAR/Tapas & Formal & Holzapfel, Waibel \par (KIT Karlsruhe) & & TFS (Typed Feature Structures) & \cite{Holzapfel2008}\\
CAST Proxies & Ubiquitous & Wyatt, Hawes, Jacobsson, Kruijff (Birmingham Univ.,
DFKI Saarbrücken) & & Amodal proxies & \cite{jacobsson2008crossmodal} \\
GSM & Structural & Mavridis, Roy \par (MIT MediaLab) & & & \cite{Mavridis2006} \\
Ke Jia Project & Formal & Chen et al. \par (Univ. of Science and Technology of China) & ASP (Answer Set Programming) & ASP & \cite{Chen2010} \\
{\sc KnowRob} & Formal & Tenorth, Beetz \par (TU München) & {\sc Prolog} & {\sc Prolog} + OWL-DL &  \cite{Tenorth2009a} \\
NKRL & Language & Zarri et al. \par (Paris Est Créteil Univ.) & NKRL & & \cite{Sabri2011} \\
%OBOC & KRS & Mendoza & & & & & \cite{Mendoza2005} \\
ORO & Formal & Lemaignan, Alami \par (LAAS-CNRS) & {\sc Java} & OWL-DL ({\sc Jena}) + {\sc Pellet} & \cite{Lemaignan2010} \\
OUR-K/OMRKF & Formal & Lim, Suh et al. \par (Hanyang Univ.) & ? & DL + Horn Clauses &  \cite{Lim2011, Suh2007} \\
PEIS KR\&R & Formal & Daoutis, Coradeshi, Loutfi, Saffiotti \par (Örebro Univ.) & {\sc C}, {\sc CycL} & CycL (1st and 2nd order logics, modal logics) & \cite{Daoutis2009} \\
%Golog & Language & Levesque (Toronto Univ.) & & {\sc Prolog} & & & \\
% & & Varadarajan, Vincze \par (TU Wien) & & & & & \cite{Varadarajan2011} \\ % -> affordances, but no implementation on a robot
% & & Kaelbling, Lozano-Pérez \par (MIT CSAIL) & & & & & \cite{Kaelbling2011} \\ % -> mostly planning under uncertainty
% & & Hertzberg (Osnabrück Univ.) \\ % -> affordances, semantic mapping
% (based on {\sc KnowRob} & & (JSK) \\

\bottomrule

\end{tabular}
\end{center}

\caption{List of surveyed systems. Categories are \emph{Formal} for systems
that have a formal underlying knowledge representation, \emph{Ubiquitous} for
systems where knowledge is fully distributed, \emph{Language} for languages
used as KRS on robots or \emph{Structural} for KRS where knowledge is
represented as special data structures.}

\label{table|surveyed-systems}
\end{table*}

%\fxfatal{Bielefeld -> could not find much\ldots}
%\fxfatal{Kollar/Tellex -> really focusing on the natural language grounding}

\subsection{ARMAR/Tapas}

{\sc Tapas} is the name of the knowledge representation system and dialogue
manager found on the ARMAR~III robot~\cite{Holzapfel2008} for the Karlsruhe
Institute of Technology.
% 
Knowledge in {\sc Tapas} exists as procedural knowledge (plans) and declarative
knowledge. The latter is split into \emph{lexical knowledge}, \emph{semantic
knowledge} and a database of identified objects (with their properties). The
\emph{lexical knowledge} contains lexical and grammatical informations about
the objects. The \emph{semantic knowledge} is organized into an ontology
relying on \emph{typed feature structures} (TFS~\cite{Carpenter1992}, a
formalism originating from the computational linguistics community, and a
superset of first-order logic).
% 
{\sc Tapas} has a strong focus on natural language grounding. It proceeds by
generating grammars from properties represented in the ontology to parse and
understand dialogue.
% 
Another focus is put on handling unknown words and objects. {\sc Tapas}
provides routines to recognize unknown entities, and propose and interactive
and iterative verbal process to categorize (including adding new categories)
those new concepts.

\paragraph{Experiments} {\sc Tapas} has been used experimentally in a kitchen
environment where naive users had to ask the robot for an object and get
information about another object.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{CAST knowledge model}
\label{sect|cast}

CAS (\emph{CoSy Architecture Schema}) Toolkit~\cite{Hawes2007} is a
comprehensive toolkit aimed at building cognitive architectures for robots
through a set of interconnected \emph{SA} (\emph{subarchitectures}). The CAS
does not expose a central knowledge base as seen in other works. It instead 
represents knowledge as unrooted \emph{proxies}. Those proxies are formally
defined in \cite{jacobsson2008crossmodal} as $p= \langle F_p, u_p \rangle$ where $F_p$ is
a set of instantiated features (like $\phi^{Colour}_{red}$) and $u_p$ a
\emph{proxies union} that form an equivalence class corresponding to one
entity.
% 
A union of proxies forms a global amodal representation of an entity that can
be explicitly shared and manipulated. Not being centralized, the knowledge
model can be qualified as \emph{ubiquitous}. Furthermore, each knowledge source 
in the CAS architecture is tightly bound to the on-line grounding process (be it
grounded in perception or in dialogue). While nothing seems to prevent it, no
{\it a priori} knowledge (including common-sense knowledge) is used.
% 
Knowledge sharing is ensured by the event mechanism of CAST -- modules can
monitor proxies for alteration by other modules. Jacobsson et al. explain how
this can be applied to reinforcement learning: The vision module creates a 
proxy for, for example, an orange object. This proxy gets monitored by a 
learning module. In parallel, the proxy can be bound to a union by the natural 
language understanding module that may add new features like \emph{"this object 
is a fruit"}. The learning module will then be called back and can add this 
new information to its model.
% 
In the presented implementation, the CAST knowledge model does no allow for
effectively representing actions or temporal information.\fxwarning{What about
reasoning? can they retrieve for example 'all proxies for colorful objects'?}

\paragraph{Knowledge Acquisition} Several techniques for knowledge acquisition
have been explored within the CAST framework. Cross-modal knowledge
fusion~\cite{Hawes2007a} is well studied, and the interaction with natural
language processing~\cite{Kruijff2010, Kruijff2010a} is a particular emphasis
of the project.
% 
In~\cite{Hawes2011}, Hawes et al. also explore \emph{curiosity} mechanisms in
the context of spatial representations with the robot \emph{Dora}.
% 
\paragraph{Experiments} CAST has been used in several experiments, including
table-top manipulation (with a focus on language understanding) and more
recently on the Dora robot~\cite{Hawes2011} for indoor exploration.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{GSM}
\label{sect|gsm}

GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
representation system primarily built to ``facilitate cross-modal
interoperability'',  especially in the context of verbal interaction with a
robot.
% 
GSM does not rely on any formal language but rather on a layered data structure
(Figure~\ref{fig|gsm}) that organizes the surrounding world into agents and
relations between agents.  Each agent (any animate or inanimate object) is
attached to a physical model (made of \emph{body parts} that have properties
like their position, color, etc.) and a mental model (which is a recursively
embedded GSM, thus allowing a sort of theory of mind).

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{gsm.pdf}

    \caption{Simplified hierarchical structure of the Grounded Situation Model,
    based on~\cite{Mavridis2006}.}

    \label{fig|gsm}
\end{figure}

Properties are represented in three layers: a stochastic representation, close
to sensory percepts, a \emph{continuous single-valued} encoding of the
stochastic model, and a discrete, categorical model.
% 
One notable feature of GSM is the \emph{bi-directionality} of the grounding
process: not only are sensor percepts abstracted into categories suitable for
human conversation, but human utterances (like ``There is a red ball in the
center of the table'') can also be turned into property descriptions. This
basically enables the knowledge representation system of the robot to
\emph{imagine} entities.
% 
GSM also features several strategies for managing time and events.
\emph{Moments} are created by storing timestamped snap-shots of the GSM, and
\emph{event classifiers} allow to define and detect events.

\paragraph{Experiments} GSM has mostly been tested on table-top manipulation
and interaction tasks (a ``conversational helping hand'' as stated by the
authors) implemented on a 7-DOF arm equipped with force feedback, cameras for blob
tracking and speech recognition (Sphinx4). Mavridis and Roy provide in addition
an in-depth analysis of the performance of GSM by means of a standard
psycholinguistic test, the \emph{Token test}~\cite{DiSimoni1978}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Ke Jia Project}
\label{sect|kejia}

The Ke Jia project~\cite{Chen2010} integrates a knowledge representation 
language with natural language processing, task planning and motion planning 
on a mobile platform.
% 
Knowledge representation relies on the \emph{Action Language C}, itself based on
\emph{Answer Set Programming} (ASP)~\cite{Gelfond2008}. These languages, that
are syntactically close to Prolog, are based on \emph{stable models} of logic
programs, and support non-monotonic reasoning. Default and non-monotonic
reasoning has been especially researched within the Ke Jia project for symbolic
task planning~\cite{Ji2011} and underspecified natural language processing.
% 
Amongst other features, the natural language processing capabilities of the
system support acquisition of new logical rules at run-time.

\paragraph{Experiments} The Ke Jia robot has been demonstrated in several tasks
involving human-robot interaction with natural language. These tasks include a
multiple \emph{pick \& carry} actions that are globally optimized, naive
physics reasoning via taught rules or more complex scenarii with the robot
delivering drinks, taking into account changing and mutually exclusive
preferences of users.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{KnowRob}
\label{sect|knowrob}

{\sc KnowRob}~\cite{Tenorth2009a,tenorth13knowrob} is an integrated knowledge 
management system originally developed at the Technische Universität München 
that has become an open-source project that is maintained by the University
of Bremen. It is build as a set of modules (Figure~\ref{fig|knowrob} that are 
organized around a core reasoning system written in Prolog. This core module 
interfaces with external modules through Java/Prolog or C/Prolog APIs.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{knowrob.pdf}

    \caption{Overview of the {\sc KnowRob} framework, taken
    from~\cite{Tenorth2011}.}

    \label{fig|knowrob}
\end{figure}

Extension modules can plug into the system to provide specialized reasoning
capabilities or interfaces to external data sources, \eg to read object
detections from the vision system. These modules operate on the level of
instances (ABox).

\paragraph{Knowledge model} {\sc KnowRob} can load OWL ontologies, and the
\emph{KnowRob-Base} ontology is provided as a common-sense ontology, with a
focus on household and kitchen domains. {\sc KnowRob} also stores and reasons 
on introspective knowledge through the \emph{Semantic Robot Description
Language}~\cite{Kunze2011} that allows to symbolically represent the robot's
capabilities.

\paragraph{Reasoning Techniques} Amongst the notable {\sc KnowRob} extensions,
{\sc ProbCog}~\cite{Jain2009} is an effort to provide probabilistic reasoning
based on Bayesian Logic Networks. The integration with naive physics reasoning 
has been studied~\cite{Kunze2011a}, and the automatic parsing of Web resources 
in natural language has been also experimented with~\cite{tenorth10webinstructions}.

\paragraph{Grounding} {\sc KnowRob} offers a mechanism called \emph{computables} 
that allows to evaluate certain predicates by calling external dedicated functions 
(for instance, the truth value of a proposition like \stmt{object1 isOn object2} 
is computed at runtime by calling a specific geometric reasoning module). In
combination with Prolog's lazy evaluation strategy, this supports a good
scalability.
Computables rely on various subsystem to evaluate, for example the robot's 
perception system~\cite{Klank2009} and \emph{semantic environment maps}~\cite{Blodow2011}
for the recognition of objects and environment.

\paragraph{Experiments} {\sc KnowRob} has been deployed in several scenarii 
on the PR2 robot and on a 2-arm custom mobile manipulator in the scope of the 
``assistive kitchen''~\cite{Beetz2008} project. These experiments included the
retrieval and automated parsing of recipes from the Web, the retrieval and 
manipulation of various kitchen tools, and cooperation between two robots.
{\sc KnowRob} has been used as knowledge base for the RoboEarth project that 
investigated the exchange of knowledge between robots via the 
Internet~\cite{waibel11roboearth}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{NKRL}
\label{sect|nkrl}

\emph{NKRL} stands for \emph{Narrative Knowledge Representation Language}.
While this language has been developed for a long time by Zarri~\cite{Zarri1997,
Zarri2008}, recent research aims at its application to robotics~\cite{Sabri2011}. 
NKRL is not {\it per-se} a knowledge representation system, it is primarily 
a language, but it is used as the representation and reasoning mechanism for 
robots by Sabri et al.

\paragraph{Knowledge representation} The NKRL language semantics are stored in
two ontologies: an ontology of concepts $\Omega$ and an ontology of events
$\Psi$. The ontology of events is made up of templates for actions and situations.
Templates are sets of predicates (MOVE, PRODUCE, RECEIVE, EXPERIENCE, BEHAVE,
OWN and EXIST) associated with thematic roles. Grounding and reasoning with NKRL
is based on template matching.

\paragraph{Experiments} The main scenario of development for NKRL-based robots
is a Smart Home that is monitoring elderly people. Knowledge acquisition
partially relies on ambient intelligence (RFID, pressure sensors in the chairs,
etc.). The scenario is still being implemented.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{OUR-K and OMRKF}
\label{sect|omrkf}

The Ontology-based Unified Robot Knowledge~\cite{Lim2011} (OUR-K) framework,
successor of the Ontology-based Multi-layered Robot Knowledge
Framework~\cite{Suh2007} (OMRKF), is a knowledge representation system based on
five inter-related \emph{classes} of knowledge (Figure~\ref{fig|omrkf}). It
proposes a layered approach to knowledge representation that allows to 
integrate the grounding process and the knowledge representation process. OUR-K's
knowledge model is implemented with a mix of Description Logics for the concept
hierarchies and Horn clauses.

\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{ourk.pdf}

    \caption{OUR-K organizes knowledge into fives \emph{classes}, each composed
    of \emph{levels}. Figure based on~\cite{Lim2011}.}

    \label{fig|omrkf}
\end{figure}

Each level of knowledge is build as three stages of ontological realization: a
\emph{meta-concept} (the level itself, like ``temporal context'', ``behavior''
or ``object feature''), a taxonomy of concepts inside this level (for instance
$cup : Object \sqsubseteq tableware : Object$) and an instantiation of the
taxonomy ($cup1 : cup$).

\paragraph{Representation} The robot's environment is in OUR-K represented as 
part of the $spaces : Model$ knowledge level as a classical three-layered map 
(metric, topological and semantic map). Objects (in $objects : Model$) are 
localized in the $spaces : Model$ through Voronoi nodes.
% 
The $Context$ knowledge class proposes an explicit statement of spatial context
(mostly geometric relations between objects), temporal context, and a more
general \emph{high-level} context inferred from spatial and temporal contexts.
% 
Finally, the $Activity$ knowledge class stores compound actions in an HTN-like
structure that is to be exploited at run-time by a planner.
\fxwarning{re-read the paper to improve the section...}

\paragraph{Experiments} Experiments conducted with OUR-K and OMRKF include
finding kitchen objects and reporting about their state to a human.  This
experiment also shows how OUR-K can deal with objects only partially matched by
their descriptor by introducing a $candidate()$ function.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{ORO}
\label{sect|oro}

\begin{figure}
    \centering
    \includegraphics[scale=0.36]{oro.pdf}

    \caption{Overview of the ORO server, taken from~\cite{Lemaignan2012}.}

    \label{fig|oro}
\end{figure}

\todo{nothing about ORO?}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{PEIS KR\&R}
\label{sect|peis-ecology}

{\sc PEIS Ecology}~\cite{Saffiotti2005} is a software \emph{ecosystem} that 
aims at the integration of autonomous robots with ambient intelligence.
\emph{PEIS} stands for \emph{Physically Embedded Intelligent System}: every
robots or intelligent device in the environment is abstracted as a PEIS. Each 
PEIS physical component is running a \emph{PEIS Kernel} instance. Communication 
between instances relies on a custom peer-to-peer communication protocol.
The PEIS architecture allows for adding new abilities through software 
components that are sharing the common \emph{tuple space} \todo{what is this 
tuple space?}.
We consider here the semantic layer~\cite{Daoutis2009}, referred as \emph{PEIS
KR\&R} which includes symbolic representation and reasoning.
% 
\fxwarning{have a look at the latest Daoutis paper!}

% More in details:
% - object identification based on viewpoint independent SIFT features
% - formalised anchoring system that explicitly match percieved attributes to predicates
% - Cyc predicates
% - ground 12 colors, based on a paper on color perception. Could be useful for us.
% - idem, they cite a paper on what spatial relations to compute
% - location of objects based on a previously provided semantic map (but not much on this semantic map)
% - two "memories": the robot memory stores the current list of percieved objects ; the archive memory stores what is not percieved anymore
% - uses directly Cyc (ie, 250 000 common sense concepts\ldots), via CycL language -> 2nd and higher order logics (quantification over predicates, functions, etc)
% Remark: using 2nd order logic (ie meta statements), it would be easy to store the knowledge of each agent
% - disambiguation in concept name by asking human to decide amongst all concepts known by Cyc
% - template based natural language
% - experiment conducted in a "smart" indoor environmement + simple robot

\paragraph{Knowledge model} The PEIS Knowledge representation system relies on
the {\sc ResearchCyc} knowledge base and the {\sc CycL} language for representing 
knowledge. The {\sc CycL} language allows to represent first-order logic 
sentences and has extensions for modal logics and higher order logics.
% 
\fxwarning{Is modal logics and higher order logics actually used in PEIS?} 
% 
As a system relying on {\sc CycL}, contexts can be expressed as
\emph{microtheories}: the truth or falsity of a set of statement depends of the
\emph{microtheory} in which these statements are evaluated.
% 
\fxwarning{OWA/CWA?}
% 
\begin{figure}
	\centering
	\includegraphics[width=0.9\columnwidth]{peis-architecture.pdf}
	\caption{The PEIS knowledge representation system, taken from~\cite{Daoutis2009}}
	\label{fig|peis-archi}
\end{figure}
% 
The PEIS KR\&R system is deeply integrated to the general PEIS Ecology
\emph{smart} environment. Figure~\ref{fig|peis-archi} gives an overview of the
interactions between PEIS knowledge processing layers.

\paragraph{Knowledge Acquisition} The primary source of knowledge in the PEIS 
system is perception. The PEIS ecosystem provides a SIFT-based object recognizer 
used in conjunction with ceiling cameras for object localization.  Other 
perceptual modalities are available, such as tracking of human motions or 
ambient environment monitoring. A template-based natural language parsing 
system can also be used to add new assertions to the knowledge base. The 
system can ask a human for help to disambiguate between concept names.

\paragraph{Anchoring} Daoutis et al. formalize the issue of anchoring as
finding a \emph{predicate grounding relation} $g \subseteq \mathcal{P} \times
\Phi \times D(\Phi)$, where $\mathcal{P}$ is a set of predicate symbols, $\Phi$
a set of percept's attributes, and $D(\Phi)$ the domain of these attributes.
% 
In the current implementation, the object category (returned by the SIFT
classifier), color, location, visibility and spatial relations (both topological 
such as \emph{at} and \emph{near} and directional such as \emph{left of} or 
\emph{behind}) are the five classes of extracted attributes.

\paragraph{Integration into the robot architecture}
\label{sect|peis-integration}

The PEIS framework offers, through the \emph{PEIS middleware}, a practical way 
to insert a new component into the shared \emph{tuple space}.  Thus, the KR\&R
module can be seamlessly integrated into the PEIS ecosystem.

\paragraph{Experiments} Experiments involving PEIS take place in a Smart Home
environment (the \emph{PEIS Home})and  explore
dialogue-based interaction with the robot about known objects.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section{Synthesis: How Do Current Systems Tackle the Knowledge Representation
Challenge?}
\label{sect|conclusion}

This section summarizes the different approaches to knowledge representation and
reasoning that are surveyed in the previous section, and attempts to identify
research perspectives.
First, the Table~\ref{table|contribution-by-systems} presents, system by system
and feature by feature, a landscape of current research, underlining the main
design choices and scientific \emph{contribution} of each of the KRS that we
have examined.
Then, we revisit the nine ``challenges'' proposed by McCarthy and Roy (see
Section~\ref{sect|evaluation-literature}) to evaluate how far we are on the 
road to natural interaction and, to slightly paraphrase McCarthy, ``human-level
robots''.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Logic formalisms, continuous world}
McCarthy underlines the necessity of relying on mathematical logic as the most
fruitful formalism for machine intelligence. Almost all of the systems we have 
surveyed rely on logical formalisms, in several cases a mix of declarative 
languages such as Description Logics and logical programming languages such as 
Prolog or Answer Set Programming.

The two exceptions (the CAST knowledge model and GSM) are however interesting: 
CAST proposes a pervasive model of knowledge that is convincing from the 
grounding perspective (but likely suboptimal for deliberative tasks), and GSM 
encodes knowledge in an amodal (continuous, geometric) model while preserving 
features that are usually specific of symbolic models (like a theory of mind 
or categorical knowledge).

The interleaving of discrete symbolic models with the continuous physics of the
world remains a major challenge, and techniques vary a lot: besides the CAST
proxies and the GSM amodal model, most systems directly extract physical
attributes of the environment to insert them in the symbolic model ({\sc Tapas},
Ke Jia, OUR-K, PEIS), thus skipping an intermediate geometric model altogether.
ORO has tight (but unidirectional) links with an intermediate geometric
representation (called {\sc Spark}), and {\sc KnowRob} adopts a top-down approach
where local geometric models are set up on demand to compute symbolic properties
like relative locations.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Management of uncertainty and approximation}

McCarthy also mentions the necessity to deal with \emph{approximate concepts
and approximate theories}, which includes representing them and reasoning with
them.
The most notable attempt at modeling uncertainty down to the knowledge
representation is {\sc ProgCog}, an extension of {\sc KnowRob}. Theoretical
and practical performances issues (including decidability) remain however
difficult to overcome.
It can be also noted that some efforts do exist to provided probabilistic
reasoning to Description Logics based systems (like the {\sc Pronto}
reasoner~\cite{Klinov2008}), but this does not appear to be a major focus in the
currently available tools. \todo{statistical relational models start to be 
used in robotics for modeling spatial relations (De Raedt ICRA 14) or for 
improving perception (Nyga ICRA 14)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Non-monotonic reasoning}

McCarthy assigns particular importance to non-monotonic reasoning. Systems 
like Ke Jia explicitly tackle this issue. A monotonic system does not 
theoretically allow for knowledge retraction during the reasoning process, 
which is an important issue in a robotic context where the world model is 
likely to change quite regularly. This is however only a practical issue
if the reasoning process has to be \emph{continuously active} during the 
robot's activity lifespan, which is rarely the case. It is often possible to
stop the reasoner, alter the knowledge, and restart the inference process on
a new domain.

Non-monotonicity can be partially dealt with using appropriate time
representation and reasoning methods (the reasoner then only takes into account
statements that are set as valid for a given moment, as it is for instance used 
in {\sc KnowRob}) or with dedicated reasoners for tasks requiring non-monotonic 
reasoning like symbolic task planning \fxwarning{mention ORO + HATP}.
Probabilistic reasoning also implicitly leads to non-monotonic
reasoning by relying on a continuous description of the state of the
world.

Non-monotonicity is however of broader significance to the knowledge
representation field, in particular for the representation of common-sense
knowledge where a \emph{default} representation is currently sorely lacking.
It is also related to what McCarthy calls \emph{elaboration tolerance}, the
ability to extend \emph{on demand} the closed domain of interpretation for a
given assertion to take new contextual knowledge into account.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Modeling of contexts}

Both Roy and McCarthy highlight the importance of formalizing and reasoning 
about contexts. Many of the KRS we have surveyed do at some point mention the 
modeling of contexts, but no consistent interpretation, let alone theory, of 
context management has clearly emerged. Several approaches for building 
contextualized knowledge have been hinted in Section~\ref{sect|features} such 
as symbolic environment interpretation, perspective taking and independent 
mental states for each agent, grounded natural language resolution, and 
self-awareness of an agent's own activities. Much remains to be done for a 
robot to actually identify its current context as well as contexts that
may be referred to.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Reasoning about time, events and actions}

While the representation of spatial roles (topology and placement with or without
perspective-taking) is well studied and KRS usually integrate with spatial
perception and reasoning components, the representation of time and its 
integration into knowledge-based systems is less uniform.
Two important tasks that require temporal reasoning are task planning and 
sequence recognition (in particular, action recognition). Systems that directly
integrate task planning (like OUR-K, {\sc KnowRob}) thus have mechanisms (like
fluents in {\sc KnowRob}) to represent time.
Other approaches include storing snapshots of the complete knowledge base 
(like in GSM) that are used to move back to past mental states of the robot 
or of another agent. Integration of these techniques with memory mechanisms 
(forgetting and reinforcement learning) remains mostly to be explored.
Regarding events, several systems (GSM, NKRL, ORO) have adopted an {\em
event-oriented} architecture where conditions (or templates) are used to
trigger decisional and execution processes.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Grounding in a multi-modal environment}

Symbol grounding remains a focus of most of the developers of cognitive
architectures for robots, and all the systems we have surveyed implement
grounding strategies. Most of the systems
adapt their grounding strategies to the sensing modality, so that grounding can
not be considered as a single, well delimited process.

Some systems (ORO, GSM) introduce an intermediate step in the grounding process
by means of an amodal model of the environment that aggregates the perceptions 
(or suppositions) in a single place. This enables geometric reasoning that 
improves the observability of the system and takes all perceptual modalities 
into account. The latter is particularly helpful for spatial perspective taking 
in order to know what humans are looking at and where objects are.

We note in addition that the development of high-level sensors is simplifying
the grounding task. The most obvious example is the human tracking system
provided with the low-cost Kinect family of devices that is now widely used
in robotics labs. Not only does this system segment and compute the poses of 
humans that enter its field of view, it also tracks them even in case they 
are partially or completely occluded. This kind of high-level sensing device 
does not remove the need of grounding one symbolic instance of a human with 
the physical human, but the task is much simpler than it used to be.

Many systems also tackle the difficult issue of natural-language processing.
The CAST middleware, in particular, has been used in the European CoSy and CogX
projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
the interest in grounding verbal interaction in the cognitive robotics community, 
and we are likely to reach important milestones in this field in the coming years.
Merging other modalities (especially back-channel communication, deictic
gestures and social gazes) also receives a lot of attention and is becoming more
and more present as symbolic knowledge available to the control layers.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{landscape}
\begin{table}\tiny
\begin{center}
\begin{tabular}{p{0.2cm}p{3.4cm}p{1.6cm}p{1.3cm}p{1.7cm}p{1.5cm}p{2cm}p{2cm}p{2cm}p{1.4cm}p{1.8cm}}
\toprule
\multicolumn{2}{c}{\bf Category}                                                     & ARMAR \cite{Holzapfel2008}& CAST \cite{Hawes2007}       & GSM \cite{Mavridis2006}     & {\sc Ke Jia} \cite{Chen2010}& {\sc KnowRob} \cite{Tenorth2009a}  & NKRL \cite{Sabri2011}                           & ORO \cite{Lemaignan2010}                      & OUR-K \cite{Lim2011}          & PEIS \cite{Daoutis2009}        \\

\midrule

\multirow{4}{*}{\turn{90}{\bf Expressiveness}}                   & Logical formalism & TFS                       & (none)                      & (none)                      & ASP                         & Prolog                             & NKRL language (FOL + 2nd order extensions)      & DL (OWL)                                      & DL + Horn clauses             & {\sc CycL}                     \\
                                                                           & OWA/CWA &                           &                             &                             & CWA                         & CWA                                &                                                 & OWA                                           &                               &                                \\
                                                              & Modeling uncertainty &                           &                             & ++ (stochastic models)      &                             & {\it+++} ({\it ProbCog} \cite{Jain2009}) & +                                         &                                               & + (\emph{candidate} entities) &                                \\
                                                                    & Meta-cognition &                           &                             &                             &                             & +                                  & + (transformation rules)                        & ++ (reification, taxonomy walking)            &                               &                                \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                             
\multirow{7}{*}{\turn{90}{\bf Representation}}                & Space Representation &                           &                             &                             &                             &                                    &                                                 & ++ ({\em perspective-aware symbolic locations} \cite{Sisbot2011})& ++         &                               \\
                                                               & Time representation &                           &                             & ++ (snapshots)              &                             & + (Allen's intervals, via \emph{computables} & ++ (event oriented)                   &                                               & +                             &                                \\
                                                                    & Actions/Events &                           &                             & ++ (events classifiers)     &                             & ++ (action recognition \cite{Beetz2010a})  & +++ (everything is an event)            &                                               & +++                           &                                \\
                                                                           & Context &                           &                             &                             &                             &                                    & ++ (template matching)                          &                                               & ++                            & ++ (microtheories)             \\
                                                                          & Modality &                           &                             & ++ (recursive model)        &                             &                                    &                                                 & ++ (Theory of Mind \cite{Warnier2012a})       &                               &                                \\
                                                                    & Self-knowledge &                           &                             &                             &                             & + (SRDF \cite{Kunze2011})          &                                                 &                                               &                               &                                \\
                                                                     & Memory models &                           &                             &                             &                             &                                    &                                                 & +                                             &                               & +                              \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                             
\multirow{9}{*}{\turn{90}{\bf Reasoning}}                   & Standard FOL reasoning &                           &                             &                             & +++                         & +++                                & + (template matching)                           & ++                                            & +                             & ++                             \\
                                            & Instantiation and structure alteration & ++                        &                             &                             & ++                          & + (dynamic instantiation)          &                                                 & ++ (TBox alteration)                          &                               &                                \\
                                                                   & Lazy evaluation &                           &                             &                             &                             & +++ (Prolog, computables)          &                                                 &                                               &                               &                                \\
                                                           & Non-monotonic reasoning &                           & ++ \cite{Hawes2011}         &                             & +++ \cite{Ji2011}           & + \cite{Tenorth2012}               &                                                 &                                               &                               &                                \\
                                                      & Presupposition accommodation &                           &                             & +++                         &                             &                                    &                                                 &                                               &                               &                                \\
                                    & Prediction, projection, diagnosis, explanation &                           &                             &                             &                             & + \cite{Tenorth2012}                                   & ++                                              & + (explanation)                               &                               &                                \\
                                                                     & Task planning &                           &                             &                             & ++ \cite{Ji2011}            & +                                  &                                                 & +++ ({\em HATP})                              & +                             & {\it++}                        \\
                                                           & Physics-based reasoning &                           &                             &                             & +                           & {\it+++} ({\em naive physics \cite{Kunze2011a}}) &                                   &                                               &                               &                                \\
%                                                                         & Learning & ++                        & ++ \cite{Jacobsson2008}     &                             &                             &                                    &                                                 &                                               &                               &                                \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                             
\multirow{5}{*}{\turn{90}{\bf Acquisition}}                         & Cross-modality & + (pointing gestures)     & ++                          & +++ (amodal model)          &                             &                                    &                                                 & ++                                            &                               & ++ ({\em ambient intelligence})\\
                                                                               & NLP & +++                       & +++ \cite{Kruijff2010a}     & +++                         & +++                         &                                    &                                                 & +++ ({\em Dialogs} \cite{Lemaignan2011a})     &                               & + (template-based)             \\
                                                                     & Web Resources &                           &                             &                             &                             & {\it+++} ({\em Web content processing \cite{tenorth10webinstructions}})&                             &                                               &                               &                                \\
                                                                         & Grounding &                           & ++                          & ++ (bidirectional)          &                             & +++ ({\em semantic maps \cite{Blodow2011, Klank2009}}) &                             & ++ ({\em amodal model} \cite{Lemaignan2012})  & +++ (bottom-up)               & +++ \cite{Loutfi2008}          \\
                                                              & Intrinsic Motivation &                           & ++ \cite{Hawes2011}         &                             &                             &                                    &                                                 &                                               &                               &                                \\
\hline                                                                                                                                                                                                                                                                                                                                                                                                             
\multirow{4}{*}{\turn{90}{\bf Integration}}     &   \ldots with sensori-motor layers &                           &                             & +                           &                             & +++ (computables, local geometric models) &                                          &                                               & +                             & ++ (tuple space)               \\
                                                      & \ldots with executive layers &                           & ++ (ubiquitous events)      & +                           &                             & ++ (language extensions) \cite{Beetz2010} &                                          & ++ (semantic events)                          & ++                            & ++ (tuple space)               \\
                                                          & Monitoring and debugging &                           &                             &                             &                             & +(Prolog tracer, visualisations...)&                                                 & + ({\em remote visualisation})                &                               &                                \\
                                                           & Performances evaluation &                           & ++ \cite{Hawes2008}         & ++ (Token test)             & +                           & + \cite{Tenorth2011}               &                                                 &                                               &                               &                                \\

\bottomrule

%\end{tabularx}
\end{tabular}
\end{center}

\caption{Design choices and main domains of contribution of current KRS (+
indicates \emph{moderate contribution}, +++ indicates \emph{major
contribution}). Italics mean that the feature is implemented as an external
module. Main references are given in the table header. When relevant,
feature-specific publications have been provided.  An empty cell means that
either the system has no specific focus on this domain or we could not find
relevant literature.}

\label{table|contribution-by-systems}
\end{table}
\end{landscape}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Common-sense}
McCarthy actually starts his list by affirming that intelligent systems must be
able to \emph{``operate successfully in the common sense informatic
situation''}. This question of \emph{common sense} is probably one of the
toughest because what \emph{common sense} means is not clear in the first place
(since, by definition, common sense is, well, common sense...). We feel,
however, that it is related to a diffuse cultural background, and one may even
claim that for a system to acquire common sense reasoning is equivalent to
solving the strong AI challenge (this is at least more or less McCarthy's opinion).

Knowledge representation systems usually follow a pragmatic approach to
common-sense knowledge by reusing existing knowledge bases, e.g. from the 
Web. While {\sc KnowRob} is the only one explicitly working on common-sense 
reasoning through physics-based reasoning, integration with Web databases and 
parsing of semi-structured Web documents, several other KRS rely on Web 
standards (OWL, OpenCyc) to represent their knowledge (Table~\ref{table|knowledge-sources} 
lists the sources of common-sense knowledge found in the surveyed systems). With the
development of initiatives like OpenMind\ref{TODO} that collect large amounts 
of common-sense facts and rules, there is a strong potential for robots to gain
common-sense knowledge from well-structured online resources in the coming
years.

%%%%%%%%%% Underlying knowledge model table %%%%%%%%
\begin{table}
\begin{center}

\begin{tabular}{lp{4cm}}
\toprule
{\bf Project} & {\bf Common-sense \par knowledge source} \\
\midrule
{\sc KnowRob} & {\sc OpenCyc}, processed web content, custom OWL-DL ontology \\
ORO & {\sc OpenCyc}, custom OWL-DL ontology \\
PEIS Ecology & {\sc ResearchCyc} \\
NKLR &  None \\
CAST Proxies &  None \\
GSM &  Predefined categories \\
OUR-K & {\it A priori} knowledge structure and axioms, custom set of instances\\
Ke Jia & None \\
ARMAR/{\sc Tapas} & Custom ontology related to the kitchen\\

\bottomrule

\end{tabular}
\end{center}
\caption{Underlying knowledge sources for each project}
\label{table|knowledge-sources}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Learning, representation of experience, introspection}

Learning at the level of the knowledge representation system has been explored
in conjunction with visual perception systems (mentioned for instance for the
CAST project~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its 
pool of facts from informations automatically extracted from the 
Web~\cite{tenorth10webinstructions}, can also be considered as a \emph{learning} 
system (as one that autonomously acquires knowledge). We do however lack a 
formal study of learning strategies and techniques at the knowledge 
representation level.

The representation and matching of past experience is a related topic. This ability
is a key step for general action recognition, and is of particular importance
for the robot to assess the state of the interaction with the human.
While we already mentioned that several systems are able to reason on past
states, we are not aware of existing implementation of algorithms to reflect on
past experiences. \todo{RACE, CRAM$_m$}

This topic is related to introspection and meta-cognition: the shift towards
explicit knowledge representation exemplified in the nine systems we have
presented has major impact on the meta-cognitive capabilities of our robots.
They can exhibit, manipulate and reason on their internal belief state (what we
will later call the \emph{cognitive observability}). We still have to discover 
all the possibilities that are open by this important cognitive ability.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Perspective-awareness}

Finally, Roy mentions \emph{the ability to take into account the human
perspective}. This cognitive ability that relates to the representation of
different modalities is explicitly advertised in ORO and GSM only. ORO
explicits the perspectives in different symbolic mental states while GSM
recursively stores models for each agent.  We think however that several 
of the other systems could easily gain support for alternative belief models, 
leading the way to becoming perspective-aware KRS.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Conclusion: Future Directions to Explore}

TDB

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{ieeetr}
\bibliography{biblio}

\end{document}